---
title: "Figure 3"
author: "Felix Wu"
date: "2026-01-21"
output:
  html_document:
    toc: true
    toc_float: true
---

Rmarkdown notebook for plotting the repeat figures (Figure 3 in draft manuscript)

## Load packages, data

Load packages

```{r loadPkg, message=FALSE}
source("../utils/genome_plots.R")

req_pkgs <- c("data.table", "stringr", "ggpubr", "patchwork", "ggpattern")
LoadPackages(req_pkgs)

SetBaseTheme(base_family = "Arial", fontpath = "~/.local/share/fonts")
```

Load repeat data. This is genome assembly repeat abundances.
```{r loadRptData}
rpt_dat <- fread("tables/REPEAT_SUMMARY.tab", header = TRUE, sep = "\t")
rpt_dat <- dcast(rpt_dat, repID + repFamily + repOrder + repSuperfamily + repClade ~ spp, value.var = c("indRepCount", "indRepBP"))
old_nms <- str_subset(names(rpt_dat), "_roe$") # Fix shortened roemeriana name
setnames(rpt_dat, old_nms, str_c(old_nms, "m"))
```

Load TE-SV data. This includes counts across all SVs in a category and 
subsamplings of SVs in the *Phlox sp.*-present category, which are required 
for bootstrap p-value calculation.


```{r loadTESVdata}
nsubsamp <- 500 # This many subsamples were taken

tesv_dat <- NULL
for (nm in c("1sp", "2sp", "3sp", "pres")) {
    tmp <- fread(str_interp("te_sv_counts/te_sv_counts.${nm}.tsv"), header = TRUE)
    if (is.null(tesv_dat)) {
        tesv_dat <- tmp
    } else {
        tesv_dat <- merge(tesv_dat, tmp, by = "repID")
    }
}

# Phlox sp.-present subsamples
for (i in 1:nsubsamp) {
    tmp <- fread(str_interp("te_sv_counts/presSubsamp/te_sv_counts.presSubsamp.${i}.tsv"),
                 header = TRUE)
    old_nms <- str_subset(names(tmp), "repID", negate = TRUE)
    new_nms <- str_c(old_nms, "N", i)
    setnames(tmp, old_nms, new_nms)
    tesv_dat <- merge(tesv_dat, tmp, by = "repID")
}
rm(tmp)
gc()

# Filter out simple repeats
tesv_dat <- tesv_dat[!str_detect(repID, "^\\(.*n$")][!str_detect(repID, "-rich$")]

# Merge with repeat data
tesv_dat <- merge(tesv_dat, rpt_dat, by = "repID", all.x = TRUE)
for (sp in c("drum", "cusp", "roem")) {
    newcol <- str_interp("${sp}Shared")
    newval <- tesv_dat[[str_interp("${sp}Pres")]] - tesv_dat[[str_interp("${sp}Uniq")]]
    tesv_dat[, eval(newcol) := newval]
}
```

Create list objects to hold plots and regression data for later joint plotting w/ patchwork
```{r plotList}
plist <- list()
mlist <- list()
```


## TE-SV comparisons

These are comparisons include:
*  TE-SV counts vs TE-SV counts. E.g., counts for SVs found in all species vs just one species.
*  Overall genomic abundance vs TE-SV counts

First load comparisons

```{r comparisonsTab}
comp_tab <- fread("te_sv_comparisons.tsv", header = TRUE, sep = "\t")
comp_tab[, NAME := str_c(YVAR, "::", XVAR)]
setkey(comp_tab, NAME)
print(comp_tab)
```

### Regression

Compute GLM for each comparison. We're using Negative Binomial with identity link due to 
count data and fixing the intercept at 0 (the origin)

```{r tesvGLM}
for (nm in comp_tab[["NAME"]]) {
    xvar <- comp_tab[nm, XVAR]
    yvar <- comp_tab[nm, YVAR]
    mlist[[nm]] <- RunNegBinom(tesv_dat, xvar = xvar, yvar = yvar,
                               link = "identity", ols_start = TRUE)
}
```

### Plotting

Plot TE-SV counts and GLM results for each comparison.

```{r plotTESV, fig.width=4, fig.height=4}
for (nm in comp_tab[["NAME"]]) {
    m <- mlist[[nm]]
    labelstr <- FormatR2PvalLab(c(m$rsq, m$pval), stacked = TRUE, pseudo = TRUE)

    xvar <- comp_tab[nm, XVAR]
    yvar <- comp_tab[nm, YVAR]

    xlab_str <- comp_tab[nm, XLAB]
    ylab_str <- comp_tab[nm, YLAB]

    pts_color <- "black"

    p <- ggplot() +
        geom_point(aes(x = .data[[xvar]], y = .data[[yvar]]), data = tesv_dat, color = pts_color, alpha = 0.5) +
        geom_ribbon(aes(x = .data[[xvar]], y = .data[[yvar]], ymin = LWR, ymax = UPR),
                    data = m$fitdat, fill = "gray", color = NA, alpha = 0.8) +
        geom_line(aes(x = .data[[xvar]], y = .data[[yvar]]), data = m$fitdat, color = "blue") +
        annotate("text", x = -Inf, y = Inf, hjust = -0.1, vjust = 1, label = labelstr, parse = TRUE, family = "Arial")

    if (str_starts(xlab_str, "bquote")) {
        p <- p + xlab(eval(parse(text = xlab_str)))
    } else {
        p <- p + xlab(xlab_str)
    }

    if (str_starts(ylab_str, "bquote")) {
        p <- p + ylab(eval(parse(text = ylab_str)))
    } else {
        p <- p + ylab(ylab_str)
    }

    plist[[nm]] <- p
    print(p)
}
```

## cuspPresSubsamp vs drumPresSubsamp bootstrap

We observe that TE counts for SVs unique to *P. drummondii* explains less variance in 
TE counts for SVs unique to *P. cuspidata* compared to TE counts for SVs that are merely 
detected in the two species (i.e., not requiring uniqueness).

To assess the likelihood of this observation, we take a bootstrap approach, subsampling
the complement of SVs present in *P. drummondii* and *P. cuspidata* to match number of SVs
found to be respectively unique and count TEs as usual. We then compute a distribution of 
(pseudo-) R^2 values in their correlations and compare them to that observed in the unique 
vs unique set.

```{r bootPresSubsampLM}
m_bootlist <- list()

for (i in 1:nsubsamp) {
    xvar <- str_c("drumPresSubsampN", i)
    yvar <- str_c("cuspPresSubsampN", i)
    m_bootlist[[i]] <- RunNegBinom(tesv_dat, xvar = xvar, yvar = yvar,
                                   link = "identity", ols_start = TRUE)
}
```

Plot the pseudo-R^2 distribution along with observed
```{r plotBootR2, fig.width=4, fig.height=3}
r2_boot <- data.table(
    RSQ = unlist(lapply(m_bootlist, function(x) x$rsq))
)

r2_obs <- mlist[["cuspUniq::drumUniq"]]$rsq
pval <- sum(r2_obs >= r2_boot$RSQ) / nsubsamp
if (pval == 0) {
    pval_lab <- str_c("italic(p) < ", FloatAsSci(1 / nsubsamp))
} else {
    pval_lab <- str_c("italic(p) == ", FloatAsSci(pval))
}

p <- ggplot(r2_boot, aes(x = RSQ)) +
    geom_density(fill = "steelblue1", color="black") +
    # geom_histogram(bins = nclass.FD(r2_boot$RSQ), fill = "steelblue1", color = "black") +
    geom_vline(aes(xintercept = r2_obs), color = "red") +
    annotate("text", x = -Inf, y = Inf, hjust = -0.1, vjust = 1,
             label = pval_lab, parse = TRUE, family = "Arial", size = 3) +
    xlab(bquote("Pseudo-" * italic(R)^2)) +
    ylab("Density")

plist[["boot_hist"]] <- p
print(p)
```

## Conceptual plot

Let's make a visual aid to help demonstrate how TE-SV counts work

### Load genotype data
```{r loadGT}
gt_dat <- fread("tables/phlox_variation_summary.FINAL.gt.tsv.gz", header = TRUE)[CHROM == "chrlg1_RagTag"]
gt_dat[, `pdrum_REF:GT` := NULL]

# Simplify column names
old_nms <- str_subset(names(gt_dat), ":GT$")
new_nms <- str_remove(old_nms, ":GT$")
gt_cols <- new_nms <- unname(sapply(new_nms, RenameWGS))
setnames(gt_dat, old_nms, new_nms)

# Filter for only complete data
rm_i <- apply(gt_dat[, gt_cols, with = FALSE], 1, function(x) any(x == "./."))
gt_dat <- gt_dat[!rm_i, ]
```

### Pick a window to plot
```{r pickExampleWindow}
# Annotate SV type
for (x in c("INS", "DEL", "DUP", "INV")) {
    gt_dat[str_detect(ID, x), SVTYPE := x]
}

# List of samples by species
samp_list <- list()
for (sp in SpeciesNames()) {
    samp_list[[sp]] <- str_subset(gt_cols, str_c("^", toupper(substr(sp, 1, 1))))
}

# 1. Compute per-row presence/absence by species -----------------------------

# presence (at least one non-0/0 genotype) in each species
gt_dat[, presence_cusp := rowSums(.SD != "0/0") > 0, .SDcols = samp_list$cusp]
gt_dat[, presence_drum := rowSums(.SD != "0/0") > 0, .SDcols = samp_list$drum]
gt_dat[, presence_roem := rowSums(.SD != "0/0") > 0, .SDcols = samp_list$roem]

# species-unique and shared SVs
gt_dat[, drum_only  := presence_drum & !presence_cusp & !presence_roem]
gt_dat[, cusp_only  := presence_cusp & !presence_drum & !presence_roem]
gt_dat[, roem_only  := presence_roem & !presence_drum & !presence_cusp]

gt_dat[, drum_shared := presence_drum & (presence_cusp | presence_roem)]
gt_dat[, cusp_shared := presence_cusp & (presence_drum | presence_roem)]

# restrict to INS / DEL only
gt_dat[, valid_type := SVTYPE %in% c("INS", "DEL")]

# convenience logicals for types
gt_dat[, is_INS := SVTYPE == "INS"]
gt_dat[, is_DEL := SVTYPE == "DEL"]

# 2. Scan windows of size w to find the first satisfying all conditions -----

# Window width
w <- 20

N <- nrow(gt_dat)
start_idx <- c()

if (N >= w) {
    for (i in 1:(N - w + 1)) {
        win <- i:(i + w - 1)

        # (1) mix of INS and DEL types only, and both types present
        cond_types_only   <- all(gt_dat$valid_type[win])
        cond_has_INS      <- any(gt_dat$is_INS[win])
        cond_has_DEL      <- any(gt_dat$is_DEL[win])

        # (2) at least one SV only in drum
        cond_drum_only    <- any(gt_dat$drum_only[win])

        # (3) at least one SV only in cusp
        cond_cusp_only    <- any(gt_dat$cusp_only[win])

        # (4) at least one SV present in drum but not unique to drum
        cond_drum_shared  <- any(gt_dat$drum_shared[win])

        # (5) at least one SV present in cusp but not unique to cusp
        cond_cusp_shared  <- any(gt_dat$cusp_shared[win])

        if (cond_types_only &&
            cond_has_INS && cond_has_DEL &&
            cond_drum_only && cond_cusp_only &&
            cond_drum_shared && cond_cusp_shared) {
            start_idx <- c(start_idx, i)
        }
    }
}
```

### Plot genotype matrix
```{r plotGtMatrix}
# Extract window
if (!is.null(start_idx)) {
    i <- start_idx[1200]
    gt_window <- gt_dat[i:(i + w - 1)]
    print(gt_window)
} else {
    warning("No good window found")
}

non_sample_cols <- c("POS", "SVTYPE")
sample_cols <- unlist(samp_list)

gt_long <- melt(
    gt_window,
    id.vars = non_sample_cols,
    measure.vars = sample_cols,
    variable.name = "sample",
    value.name   = "genotype"
)

# factor for samples in the exact column order of gt_window
gt_long[, sample := factor(sample, levels = sample_cols)]

# For variants: use POS as factor, ordered by POS.
gt_long[, POS_factor := factor(POS, levels = unique(POS[order(POS)]))]

# Colors
drum_pink <- unname(Load4SpPal("2drum")["drumPink"])
geno_colors <- colorRampPalette(c("white", drum_pink))(3)
names(geno_colors) <- c("0/0", "0/1", "1/1")

# For patterning
gt_long[, nonzero := genotype != "0/0"]

# Plot
p <- ggplot(gt_long, aes(x = sample, y = POS_factor,
                         fill = genotype, pattern = SVTYPE, pattern_alpha = nonzero)) +
    geom_tile_pattern(
        color          = "grey40",     # cell borders (optional)
        pattern_fill   = "black",
        pattern_colour = "black",
        pattern_angle  = 45,
        pattern_density = 0.02,
        pattern_spacing = 0.02,
        pattern_key_scale_factor = 0.6
    ) +
    # Annotation lines
    annotate("path", x = c(4.5,  4.5), y = c(0, nrow(gt_window)) + 0.5, color = "black", size = 1.5) +
    annotate("path", x = c(8.5,  8.5), y = c(0, nrow(gt_window)) + 0.5, color = "black", size = 1.5) +
    annotate("path", x = c(0.5,  4.5) + c(0.05, -0.05), y = c(0.3, 0.3), color = Load4SpPal()["cusp"], size = 2) +
    annotate("path", x = c(4.5,  8.5) + c(0.05, -0.05), y = c(0.3, 0.3), color = Load4SpPal()["drum"], size = 2) +
    annotate("path", x = c(8.5, 11.5) + c(0.05, -0.05), y = c(0.3, 0.3), color = Load4SpPal()["roem"], size = 2) +
    # Legends
    scale_fill_manual(values = geno_colors, name = "Genotype") +
    scale_pattern_manual(values = c("INS" = "stripe", "DEL" = "none"),
                         name = "SV type") +
    scale_pattern_alpha_manual(values = c(`FALSE` = 0, `TRUE` = 1),
                               guide  = "none") +
    scale_y_discrete(limits = rev(levels(gt_long$POS_factor)), name = "Chr1 position") +
    scale_x_discrete(name = "Sample") +
    coord_fixed() +  # make cells square
    theme(
        axis.text.x  = element_text(size = 9, angle = 45, hjust = 1, vjust = 1, family = "Arial"),
        axis.text.y  = element_text(size = 8, family = "Arial"),
        axis.line.x = element_blank(),
        axis.line.y = element_blank(),
        panel.grid   = element_blank(),
        axis.ticks   = element_line(color = "black"),
        legend.position = "left"
    )

print(p)
plist[["gt_matrix"]] <- p
```

```{r plotSpeciesMatrix}
species_levels <- names(samp_list)

species_long <- rbindlist(list(
    gt_window[, .(POS,
                  species = "cusp",
                  present = presence_cusp,
                  unique  = cusp_only)],
    gt_window[, .(POS,
                  species = "drum",
                  present = presence_drum,
                  unique  = drum_only)],
    gt_window[, .(POS,
                  species = "roem",
                  present = presence_roem,
                  unique  = roem_only)]
))

# Order rows and columns
species_long[, species := factor(species, levels = species_levels)]
species_long[, POS_factor := factor(POS, levels = unique(POS[order(POS)]))]

# Fill colors
species_colors <- Load4SpPal()[species_levels]

# We'll use presence to decide whether to fill with the species color or not (NA = white)
species_long[, fill_color := NA_character_]
species_long[, fill_color := ifelse(present, as.character(species), NA)]
species_long[, fill_color := factor(fill_color, levels = species_levels)]

# Plot species matrix
p <- ggplot(species_long, aes(x = species, y = POS_factor)) +
    # Tiles: colored by species if present, else white
    geom_tile(aes(fill = fill_color), color = "grey40") +
    # Black point if unique in that species
    geom_point(data = species_long[unique == TRUE], color = "black", size  = 1.8) +
    scale_fill_manual(
        values = species_colors,
        na.value = "white",
        breaks = species_levels,
        labels = ConvertSpeciesNames(species_levels, "short", "G_species"),
        name   = "Species\n(present)"
    ) +
    # Annotation path same as for gt matrix plot to make sizes the same
    annotate("path", x = c(0,  0), y = c(0.3, 0.3), color = "#FFFFFF00", size = 2) +
    scale_y_discrete(limits = rev(levels(species_long$POS_factor))) +
    coord_fixed() +  # make cells square
    theme(
        panel.grid  = element_blank(),
        # axis.text.x = element_text(size = 9, angle = 45, hjust = 1, vjust = 1),
        axis.text = element_blank(),
        axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        # axis.ticks.x  = element_line(color = "black"),
        axis.ticks  = element_blank(),
        axis.line.x = element_blank(),
        axis.line.y = element_blank(),
        legend.text = element_text(family = "Arial", face = "italic"),
        legend.position = "left"
    )

print(p)
plist[["sp_matrix"]] <- p
```

## Joint plots

Use patchwork syntax to display key plots together

```{r plotMatricesJoint, fig.width=10, fig.height=10}
p <- plist[["gt_matrix"]] + plist[["sp_matrix"]] +
    plot_layout(ncol = 2, axes = "collect", guides = "collect") &
    theme(legend.position = "left", legend.box = "vertical")

ggsave(p, file = "figure3.part1.pdf",
       device = cairo_pdf,
       family = "Arial",
       width = 8,
       height = 6)
print(p)
```

```{r plotTeSvJoint, fig.width=10, fig.height=10}
p1 <- plist[["indRepCount_drum::drumPres"]]
p2 <- plist[["cuspPres::drumPres"]]
p3 <- plist[["drumPres::drumUniq"]]
p4 <- plist[["cuspUniq::drumUniq"]]
p5 <- plist[["boot_hist"]] +
    theme(plot.background = element_blank(),
          panel.background = element_blank(),
          axis.title = element_text(size = 7),
          axis.text = element_text(size = 6))

p <- p1 + p2 + p3 + p4 +
    inset_element(p5, left = 0.62, bottom = 0, right = 1, top = 0.35) +
    plot_layout(ncol = 2)
ggsave(p, file = "figure3.part2.pdf",
       device = cairo_pdf,
       family = "Arial",
       width = 10,
       height = 10)
print(p)
```
