---
title: "Figure 1"
date: "2025-04-15"
output: 
    html_document:
        toc: true
        toc_float: true
author: Felix L. Wu
---


# Setup

## Load packages

```{r loadPkg, message = FALSE}
source("../utils/genome_plots.R")

# Install packages
req_pkgs <- c(
    "data.table",
    "stringr",
    "circlize",
    "shades",
    "ggpubr",
    "ComplexHeatmap",
    "pheatmap",
    "RColorBrewer",
    "patchwork",
    "ggrepel"
)
LoadPackages(req_pkgs)

SetBaseTheme(base_family = "Arial", fontpath = "../utils/fonts")
par(family = "Arial")
```

## Plot setup
```{r plotSettings}
include_gaps <- FALSE # Include gaps in circle proportion plot?
```

# Load data

## Genome sizes
```{r loadGnmSize}
ref <- fread(
    cmd = "grep '^chr' results/ref/phlox_flye.v1.0.FINAL.fasta.fai",
    header = FALSE,
    select = 1:2,
    col.names = c("CHROM", "SIZE")
)
ref[, CHROM := str_replace(CHROM, "chrlg", "chr")]
```

## Features 
```{r loadFeats}
prefix <- "results/coverage/disjoint.w_1Mb"
midfixes <- c("gaps", "exon", "intron", "repeats.LTR", "repeats.other")

gene_str <- "gene"
intergenic_str <- "intergenic"

feat_names <- c(
    "gaps" = "gaps",
    "exon" = "exon",
    "intron" = "intron",
    "repeats.LTR" = "LTR",
    "repeats.other" = "other repeats"
)

feat_levs <- list()
if (include_gaps) {
    feat_levs[["ei"]] <- c("exon", "intron", "LTR",
                           "other repeats", "gaps", intergenic_str)
    feat_levs[["gn"]] <- c(gene_str, "LTR",
                           "other repeats", "gaps", intergenic_str)
} else {
    feat_levs[["ei"]] <- c("exon", "intron",
                           "LTR", "other repeats", intergenic_str)
    feat_levs[["gn"]] <- c(gene_str, "LTR",
                           "other repeats", intergenic_str)
}

dat <- list()
dat[["ei"]] <- list()
for (mf in midfixes) {
    fn <- str_interp("${prefix}.${mf}.bed")
    dat[["ei"]][[mf]] <- fread(
        fn,
        header = FALSE,
        select = c(1:3, 7, 8),
        col.names = c(
            "CHROM",
            "START",
            "END",
            "NBASE",
            "NFEAT"
        )
    )
    dat[["ei"]][[mf]][, FEATURE := feat_names[mf]]
}
dat[["ei"]] <- rbindlist(dat[["ei"]])
dat[["ei"]][str_starts(CHROM, "chr"), CHROM := str_replace(CHROM, "chrlg", "chr")]

tmp <- dat[["ei"]][
    FEATURE %in% c("exon", "intron"), 
    .(NBASE = sum(NBASE), NFEAT = sum(NFEAT), FEATURE = "gene"), 
    by = .(CHROM, START, END)]
dat[["gn"]] <- rbindlist(list(dat[["ei"]][!(FEATURE %in% c("exon", "intron"))], tmp))

dat[["ei"]][, INDEX := str_c(CHROM, START, sep = "_")]
dat[["gn"]][, INDEX := str_c(CHROM, START, sep = "_")]
```

Process Features

```{r procFeats}
for (nm in names(dat)) {
    # Remove windows with excessively high gap proportion (> 90%)
    rm_i <- dat[[nm]][str_starts(CHROM, "chr")][FEATURE == "gaps"][(NBASE /(END - START) > 0.9), INDEX]
    dat[[nm]] <- dat[[nm]][!(INDEX %in% rm_i)]

    if (include_gaps) {
        dat[[nm]][, WSIZE := END - START]
        dat[[nm]][, PROP_BP := NBASE / WSIZE]
        tmp <- dat[[nm]][,
                         .(PROP_BP = 1 - sum(PROP_BP),
                           NBASE = NA,
                           NFEAT = NA,
                           FEATURE = intergenic_str),
                         by = .(INDEX,
                                CHROM,
                                START,
                                END,
                                WSIZE)]
        dat[[nm]] <- rbindlist(list(dat[[nm]], tmp), use.names = TRUE)
    } else {
        dat[[nm]] <- merge(dat[[nm]][FEATURE != "gaps"], dat[[nm]][FEATURE == "gaps", .(INDEX, NGAP = NBASE)], by = "INDEX")
        dat[[nm]][, WSIZE := END - START - NGAP]
        dat[[nm]][, PROP_BP := NBASE / WSIZE]
        tmp <- dat[[nm]][, .(PROP_BP = 1 - sum(PROP_BP), NBASE = NA, NFEAT = NA, NGAP = NGAP[1], FEATURE = intergenic_str), by = .(INDEX, CHROM, START, END, WSIZE)]
        dat[[nm]] <- rbindlist(list(dat[[nm]], tmp), use.names = TRUE)
    }
    dat[[nm]][, FEATURE := factor(FEATURE, levels = feat_levs[[nm]])]
    setorder(dat[[nm]], CHROM, START, FEATURE)
}

```

## Pixy data
```{r loadPixy}
pixy_info <- list()
pixy_info[["wstr"]] <- "1Mb"
pixy_info[["min_dp"]] <- 5

pixy_info[["data"]] <- list()
for (x in c("dxy", "fst", "pi")) {
    fn <- str_interp("pixy_stats/merged.ALL.DP_${pixy_info$min_dp}.${pixy_info$wstr}_${x}.txt.gz")
    pixy_info[["data"]][[x]] <- fread(fn, header = TRUE, sep = "\t")
    setnames(pixy_info[["data"]][[x]], "chromosome", "CHROM")
    pixy_info[["data"]][[x]][, CHROM := str_replace(CHROM, "chrlg", "chr")]
    pixy_info[["data"]][[x]][, COORD := (window_pos_1 - 1) / 1e6]
}
```

## SVs
```{r loadSVs}
svs <- list()
for (svtype in c("INS", "DEL")) {
    for (sp in SpeciesNames()) {
        fn <- str_interp("results/coverage/SV_${svtype}.${sp}.w_5Mb.bed")
        svs[[fn]] <- fread(
            cmd = str_c("grep '^chr' ", fn),
            header = FALSE,
            select = c(1:3,8),
            col.names = c("CHROM", "START", "END", "COUNT"))
        svs[[fn]][, SVTYPE := svtype]
        svs[[fn]][, SPECIES := sp]
    }
}
svs <- rbindlist(svs)
svs[, CHROM := str_replace(CHROM, "chrlg", "chr")]
svs[, COORD := (START) / 1e6]

# SV raw genotypes
sv_gts <- fread("results/annotations/SV.query.tsv",
                header = TRUE,
                na.strings = c(".", "./."))
new_nms <- names(sv_gts)[]
new_nms[1] <- str_remove(new_nms[1], "^#\\s*")
new_nms <- str_remove(new_nms, "^\\[[0-9]+\\]")
new_nms <- str_remove(new_nms, ":GT$")
setnames(sv_gts, names(sv_gts), new_nms)
```
# Circle figure

## Init
```{r circInit, fig.show = "hide"}
tr_h <- 0.1 # A "unit" of track height to work off of

circos.clear()
circos.par( points.overflow.warning = FALSE,
            track.height = tr_h, 
            start.degree = 89,
            gap.degree   = c(rep(1.5, 6), 9),
            cell.padding = c(0, 0, 0, 0), # v, <, ^, > -- default: 0.02, 1, 0.02, 1
            track.margin = c(0.01, 0.01)) # v, ^ -- default 0.01

init_d <- ref[, .(COORD = SIZE/1e6), by = CHROM]
init_m <- cbind(rep(0, nrow(init_d)), init_d$COORD)
rownames(init_m) <- init_d$CHROM

circos.initialize(rownames(init_m), xlim = init_m)

lgd <- list()

ylab_cex <- 1.3
ylab_cex_step <- 0.2

p_circ <- recordPlot()
```

## Track 1

Feature data

```{r circTrack1, fig.show = "hide"}
replayPlot(p_circ)
trN <- 1 # Track number

plt_d <- dat[["gn"]][str_starts(CHROM, "chr"), .(CHROM, X = START/1e6, Y = cumsum(PROP_BP), FEATURE), by = INDEX]
plt_d[, INDEX := NULL]
plt_d[, CHROM := factor(CHROM, levels = str_c("chr", 1:7))]

cur_ylim <- c(0, 1)
cur_yticks <- list('raw' = seq(0, 1, by = 0.25))
cur_yticks[["lab"]] <- sprintf("%0.2f", cur_yticks$raw)

FourColorFunc <- function() {
    return(rev(c(colorRampPalette(c("#FFFFFF", Load4SpPal("2drum")["drumPink"], "#000000"))(14)[c(2,6,10)], "#BBBBBB")))
}

FiveColorFunc <- function() {
    return(rev(c(colorRampPalette(c("#FFFFFF", Load4SpPal("2drum")["drumPink"], "#000000"))(14)[c(2,5,8,11)], "#B8B8B8")))
}
if (include_gaps) {
    # tr1_colors <- c("#999999", unname(Load4SpPal("2drum")[c(4,2,5,3)]))
    tr1_colors <- FiveColorFunc()
    # [1] "#B8B8B8" "#46172F" "#8D2F5E" "#C081A0" "#EFDFE7"
    # [1] "#B8B8B8" "#3C3C3E" "#606062" "#9898A0" "#E2E2E5"
} else {
    # tr1_colors <- c("#999999", unname(Load4SpPal()[c(2,4,3)]))
    tr1_colors <- FourColorFunc()
    # [1] "#BBBBBB" "#5E1F3E" "#B06289" "#EFDFE7"
    # [1] "#BBBBBB" "#474749" "#82828A" "#E2E2E5"
}
feat_order <- rev(levels(plt_d$FEATURE))
names(tr1_colors) <- feat_order


circos.track(   plt_d$CHROM, y = plt_d$Y, ylim = cur_ylim, bg.border = "black", track.height = tr_h * 2,
                panel.fun = function(x, y) {
                        circos.text(CELL_META$xcenter, 
                                    CELL_META$cell.ylim[2] + mm_y(3.7), 
                                    CELL_META$sector.index)
                        circos.axis(labels.cex = 1.0)
                    }
            )

chrom_levels <- plt_d[, levels(CHROM)]
for (i in seq_along(chrom_levels)) {
    chrom <- chrom_levels[i]
    for (feat in feat_order) {
        tmp_d <- plt_d[FEATURE == feat][CHROM == chrom]
        circos.lines(x = tmp_d$X, y = tmp_d$Y, col = tr1_colors[feat], area = TRUE, track.index = trN, sector.index = chrom, border = NA)
    }
}

circos.yaxis(   labels.cex = ylab_cex, sector.index = "chr1", side = "left", tick.length = convert_x(0.5, "mm", "chr1", track.index = trN),
                at = cur_yticks$raw, labels = cur_yticks$lab, labels.niceFacing = TRUE)
ylab_cex <- ylab_cex - ylab_cex_step

# Legend
lgd[["feat"]] <- Legend(at = names(tr1_colors), type = "grid", 
    legend_gp = gpar(fill = unname(tr1_colors)), title_position = "topleft", 
    title = "Feature occupancy")

p_circ <- recordPlot()
```

## Tracks 2-4
Pi estimates in drummondii, fst vs roem, fst vs cusp

```{r circTrack2, warnings = FALSE, messages = FALSE, fig.show = "hide"}
replayPlot(p_circ)

MakeYticks <- function(y_max, ndiv = 2) {
    # Assumes min Y is 0, Y >= 0.01
    ystep <- floor(y_max * 100 / ndiv) / 100
    return(seq(0, ystep*ndiv, by = ystep))
}

target_dark <- "#999999"

track_nums <- list(drum = 2:3, roem = 4:5, cusp = 6:7)
# names(track_nums) <- c("drum", "roem", "cusp")
for (sp in names(track_nums)) {
    trN <- track_nums[[sp]] # Track number
    use_color <- Load4SpPal()[sp]
    
    if (sp == "drum") {
        plt_d <- pixy_info$data$pi[pop == sp][!is.na(avg_pi), .(Y = mean(avg_pi)), by = .(CHROM, COORD = as.character(COORD))]
    } else {
        plt_d <- pixy_info$data$fst[(pop1 == "drum" & pop2 == sp) | (pop1 == sp & pop2 == "drum")][!is.na(avg_wc_fst),
            .(Y = mean(avg_wc_fst)), by = .(CHROM, COORD = as.character(COORD))]
    }
    plt_d[, X := as.numeric(COORD)]

    cur_ylim <- c(0, max(plt_d$Y))
    print(cur_ylim)
    cur_yticks <- list("raw" = MakeYticks(cur_ylim[2]))
    cur_yticks[["lab"]] <- sprintf("%0.2f", cur_yticks$raw)
    cur_yticks[["minor"]] <- lapply(cur_yticks["raw"],
                                    function(x) sort(c(x, (x[-length(x)] + x[-1])/2)))[[1]]
    cur_yticks[["minor"]] <- cur_yticks$minor[-1]

    sv_d <- svs[SPECIES == sp,
                .(COUNT = sum(COUNT)),
                by = .(CHROM, COORD = as.character(COORD))]
    sv_d[, X := as.numeric(COORD)]
    sv_d[, Y := COUNT / max(COUNT)]

    col_ramp <- colorRamp2(0:1, colors = c("#FFFFFF", BrightMatch(use_color, target_dark)))

    circos.track(
        sv_d$CHROM,
        x = sv_d$X,
        y = sv_d$Y,
        ylim = c(0, 1),
        bg.border = TRUE,
        track.height = tr_h/5,
        track.margin = c(0, 0.01),
        panel.fun = function(x, y) {
            for (i in seq_along(x)) {
                circos.rect(x[i],
                            CELL_META$ylim[1],
                            x[i]+5,
                            CELL_META$ylim[2],
                            col = col_ramp(y[i]),
                            border = FALSE)
            }}
    )

    circos.track(
        unique(plt_d$CHROM),
        ylim = cur_ylim,
        bg.border = NA,
        track.margin = c(0.01, 0),
        panel.fun = function(x, y) {
            circos.axis(h = "bottom", major.tick = FALSE, labels = FALSE)
        }
    )
    for (chrom in plt_d[, unique(CHROM)]) {
        for (gridline in cur_yticks$minor) {
            circos.lines(plt_d[CHROM == chrom, range(X)], rep(gridline, 2), col = "#888888", track.index = trN[2], sector.index = chrom, lty = 2, lwd = 0.5) # Axis line
        }
        circos.lines(plt_d[CHROM == chrom, X], plt_d[CHROM == chrom, Y], col = use_color, track.index = trN[2], sector.index = chrom) # Data
        # circos.lines( sv_d[CHROM == chrom, X],  sv_d[CHROM == chrom, Y], col = use_colors[2], track.index = trN, sector.index = chrom)
    }

    circos.yaxis(labels.cex = ylab_cex, sector.index = "chr1", side = "left", tick.length = convert_x(0.5, "mm", "chr1", track.index = trN[2]),
                at = cur_yticks$raw, labels = cur_yticks$lab, labels.niceFacing = TRUE)
    ylab_cex <- ylab_cex - ylab_cex_step

    # Legend
        
}
col_ramp <- colorRamp2(0:1, colors = c("#FFFFFF", target_dark))
lgd[["svs"]] <- Legend(at = seq(0, 1, by = 0.25), col_fun = col_ramp,
    title_position = "topleft", title = "Normalized SV density")

lgd[["pi_fst"]] <- Legend(at = c("Nt diversity (P. drum)", "Fst (P. roem vs. P. drum)", "Fst (P. cusp vs. P. drum)"),
    type = "lines", legend_gp = gpar(col = Load4SpPal()[names(track_nums)]), 
    title_position = "topleft", 
    title = "Genetic variation")

p_circ <- recordPlot()
```

## Legend
```{r circLegend, warnings = FALSE, messages = FALSE, fig.show = "hide"}
replayPlot(p_circ)

lgd_out = packLegend(list = lgd, direction = "horizontal")
draw(lgd_out, x = unit(4, "mm"), y = unit(4, "mm"), just = c("left", "bottom"))

p_circ <- recordPlot()
```

```{r circShow, fig.width = 13.5, fig.height = 13}
replayPlot(p_circ)
```

## Save circle figure as pdf
```{r circSave}
pdf(file = "figure1.circle.pdf", width = 13.5, height = 13)
replayPlot(p_circ)
dev.off()
```

# TE and SV figures

## Load data
```{r loadRptData}
rpt_dat <- fread("../repeat_figures/tables/REPEAT_SUMMARY.tab", header = TRUE, sep = "\t")
rpt_dat <- dcast(rpt_dat, repID + repFamily + repOrder + repSuperfamily + repClade ~ spp, value.var = c("indRepCount", "indRepBP"))

# Replace "Copia" and "Gypsy" with modern naming
for (cn in c("repFamily", "repOrder", "repSuperfamily", "repClade")) {
    rpt_dat[get(cn) == "LTR/Copia", eval(cn) := "LTR/Ty1"]
    rpt_dat[get(cn) == "LTR/Gypsy", eval(cn) := "LTR/Ty3"]
}
```



## Plotting functions

```{r plotPrep}
PrepRptPlotData <- function(
    u = "indRepBP",
    ignore_rpts = c("Simple_repeat",
                    "Low_complexity",
                    "Satellite",
                    "Unknown",
                    "rRNA",
                    "snRNA",
                    "tRNA"),
    combine_LTR = TRUE,
    ignore_low_occup_n = 4
) {
    if (!(u %in% c("indRepBP", "indRepCount"))) {
        stop(str_interp("Unit ${u} not recognized"))
    }

    cnames <- c()
    for (sp in c("cusp", "drum", "roe", "pilo")) {
        cnames[sp] <- str_interp("${u}_${sp}")
    }

    plt_d <- rpt_dat[!(repFamily %in% ignore_rpts),
                     .(cusp = sum(as.numeric(get(cnames["cusp"])),
                                  na.rm = TRUE),
                       drum = sum(as.numeric(get(cnames["drum"])),
                                  na.rm = TRUE),
                       roem = sum(as.numeric(get(cnames["roe"])),
                                  na.rm = TRUE),
                       pilo = sum(as.numeric(get(cnames["pilo"])),
                                  na.rm = TRUE)),
                     by = repFamily]


    # Sum LTR and LTR/Unknown
    if (combine_LTR) {
        tmp <- plt_d[repFamily == "LTR", ]

        for (sp in names(plt_d)[-1]) {
            plt_d[repFamily == "LTR/Unknown", eval(sp) := get(sp) + tmp[[sp]]]
        }
        plt_d <- plt_d[repFamily != "LTR"]
    }

    # Check that lowest occupancy families are all the same
    if (ignore_low_occup_n > 0) {
        ii <- 1:ignore_low_occup_n
        sp <- names(plt_d)[2]
        low_rpt_fams <- plt_d[order(get(sp)), repFamily][ii]
        for (sp in names(plt_d)[-(1:2)]) {
            tmp <- plt_d[order(get(sp)), repFamily][ii]
            if (length(setdiff(low_rpt_fams, tmp)) + length(setdiff(tmp, low_rpt_fams)) > 0) {
                stop("Not all lowest occupancy families are the same across species")
            }
        }
        plt_d <- plt_d[!(repFamily %in% low_rpt_fams)]
    }
    return(plt_d)
}

nbases <- c("cusp" = 5159793837,
            "drum" = 6290703384,
            "roem" = 7002483497,
            "pilo" = 6890161404)

ryb_pal <- colorRampPalette(rev(brewer.pal(n = 11, name = "RdYlBu")))
```


## Plot repeats as pie chart

### Repeats only

```{r plotRawPie, fig.width = 4, fig.height = 4}
plt_d <- PrepRptPlotData()

nr <- nrow(plt_d)
use_col <- get_palette("d3", k = min(10, nr))

for (sp in names(plt_d)[-1]) {
    pie(plt_d[[sp]], labels = plt_d$repFamily, main = sp, col = use_col)
}
```

### Plot drummondii repeats along with gene content

```{r plotDrumPie, fig.width = 6, fig.height = 6}
plt_d <- PrepRptPlotData(
    ignore_rpts = c(),
    ignore_low_occup_n = 0
)[, .(repFamily, drum)]

# Combine DNA, LINE elements
for (fam in c("DNA", "LINE")) {
    sub_d <- plt_d[str_starts(repFamily, fam)]
    cur_sum <- sub_d[, sum(drum)]
    plt_d <- rbind(
        plt_d[!(repFamily %in% sub_d[["repFamily"]])], 
        data.table(repFamily = fam, drum = cur_sum)
    )
}

# LTR/Other
sub_d <- plt_d[str_starts(repFamily, "LTR/") & !(repFamily %in% c("LTR/Ty1", "LTR/Ty3", "LTR/Unknown"))]
cur_sum <- sub_d[, sum(drum)]
plt_d <- rbind(
    plt_d[!(repFamily %in% sub_d[["repFamily"]])],
    data.table(repFamily = "LTR/Other", drum = cur_sum)
)

# RNA
sub_d <- plt_d[str_ends(repFamily, "RNA")]
cur_sum <- sub_d[, sum(drum)]
plt_d <- rbind(
    plt_d[!(repFamily %in% sub_d[["repFamily"]])],
    data.table(repFamily = "RNA", drum = cur_sum)
)

setorderv(plt_d, "drum", order = -1L)

# Combine with overall proportions of features
all_feat <- dat[["ei"]][,
                        .(PROP = sum(PROP_BP * WSIZE) / sum(WSIZE)),
                        by = FEATURE]
tmp <- all_feat[, PROP]
names(tmp) <- all_feat[, FEATURE]
all_feat <- tmp

blist <- list(
    data.table(repFamily = c("exon", "intron"),
               FEATURE = c("exon", "intron"),
               count = all_feat[c("exon", "intron")]),
    data.table(repFamily = intergenic_str,
               FEATURE = intergenic_str,
               count = all_feat[intergenic_str]),
    plt_d[str_starts(repFamily, "LTR"),
          .(repFamily,
            FEATURE = "LTR",
            count = all_feat["LTR"]* drum / sum(drum))],
    plt_d[!str_starts(repFamily, "LTR"),
          .(repFamily,
            FEATURE = "other repeats",
            count = all_feat["other repeats"]* drum / sum(drum))]
)
plt_d <- rbindlist(blist)
plt_d[repFamily %in% c("exon", "intron"), FEATURE := "gene"]
plt_d[, FEATURE := factor(FEATURE, levels = feat_levs$gn)]
setorderv(plt_d, c("FEATURE", "count"), order = c(1L, -1L))
plt_d[, repFamily := factor(repFamily, levels = unique(repFamily))]

radii <- c(2, 1)

inner_d <- plt_d[, .(count = sum(count)), by = FEATURE]
inner_c <- rev(FourColorFunc())
names(inner_c) <- levels(inner_d$FEATURE)

gene_c <- colorRampPalette(inner_c[1:2])(3)[1:2]
names(gene_c) <- c("exon", "intron")

nms <- plt_d[FEATURE == "LTR", as.character(unique(repFamily))]
ltr_c <- colorRampPalette(c(inner_c["LTR"], "#928CE0"))(length(nms))
names(ltr_c) <- nms

nms <- plt_d[FEATURE == "other repeats", as.character(unique(repFamily))]
oth_c <- colorRampPalette(c(inner_c["other repeats"],
                          Load4SpPal()["roem"]))(length(nms))
names(oth_c) <- nms

use_colors <- c(gene_c,
                inner_c,
                ltr_c,
                oth_c)

inner_d[, X := radii[1] / 2]
inner_d[, X_LAB := X]
inner_d[FEATURE == gene_str, X_LAB := X_LAB + 0.63]
inner_d[FEATURE == intergenic_str, X_LAB := X_LAB + 0.55]
# inner_d[FEATURE == "exon", X_LAB := X_LAB + 0.45]
# inner_d[FEATURE == "intron", X_LAB := X_LAB - 0.1]
inner_d[.N:1, Y := cumsum(count)]
inner_d[.N:1, Y_LAB := (Y + c(0, cumsum(count[-.N]))) / 2]
inner_d[, LABEL := str_c(FEATURE, sprintf("\n(%0.1f%%)", count * 100))]
inner_d[, LAB_C := "black"]
inner_d[FEATURE == "other repeats", LAB_C := "white"]

plt_d[, X := radii[1] + radii[2] / 2]
plt_d[, X_LAB := sum(radii) + 0.25]
plt_d[FEATURE == "gene", X_LAB := radii[1] + 0.25]
plt_d[  FEATURE == "other repeats" & repFamily != "Unknown", 
        X_LAB := X_LAB + 0.1]
plt_d[.N:1, Y := cumsum(count)]
plt_d[.N:1, Y_LAB := (Y + c(0, cumsum(count[-.N]))) / 2]
plt_d[  FEATURE == "other repeats" & repFamily != "Unknown", 
        Y_LAB := median(range(Y_LAB)) + seq(-0.028, 0.028, length.out = .N)[order(Y)]]
plt_d[, VJUST := 0.5  + 0.5 * sin(2 * pi * Y_LAB)]
plt_d[, HJUST := 0.5  - 0.5 * cos(2 * pi * Y_LAB)]
plt_d[, ANGLE := 0]
plt_d[FEATURE == "other repeats" & repFamily != "Unknown",
      ANGLE := Y_LAB * 360]
plt_d[, LABEL := str_replace(as.character(repFamily), "_", " ")]
plt_d[FEATURE == "gene",
      LABEL := str_c(repFamily, sprintf(" (%0.1f%%)", count * 100))]

p <- ggplot() + 
    geom_bar(aes(x = X, y = count, fill = repFamily), data = plt_d, 
        stat = "identity", position = "stack", width = radii[2], color = NA) +
    geom_bar(aes(x = X, y = count, fill = FEATURE), data = inner_d, 
        stat = "identity", position = "stack", width = radii[1], color = NA) +
    scale_fill_manual(values = use_colors, guide = "none") +
    geom_vline(xintercept = radii[1], color = "white")

for (i in 1:nrow(inner_d)) {
    p <- p + annotate(geom = "line", x = c(0, sum(radii)), y = inner_d[i, rep(Y, 2)], color = "white")
}
p <- p + 
    geom_text(aes(x = X_LAB,
                  y = Y_LAB,
                  label = LABEL),
              data = inner_d[LAB_C == "black"],
              size = 2.8,
              hjust = 0.5,
              vjust = 0.5,
              color = "black") +
    geom_text(aes(x = X_LAB,
                  y = Y_LAB,
                  label = LABEL),
              data = inner_d[LAB_C == "white"], 
              size = 2.8,
              hjust = 0.5,
              vjust = 0.5,
              color = "white") +
    geom_text(aes(x = X_LAB,
                  y = Y_LAB,
                  label = LABEL,
                  hjust = HJUST,
                  vjust = VJUST,
                  angle = ANGLE,
                  color = repFamily),
              data = plt_d[!(FEATURE %in% c(intergenic_str, "gene"))],
              size = 3) +
    scale_color_manual(values = use_colors, guide = "none") +
    geom_text(aes(x = X_LAB,
                  y = Y_LAB,
                  label = LABEL,
                  hjust = HJUST,
                  vjust = VJUST,
                  angle = ANGLE),
              data = plt_d[FEATURE %in% "gene"],
              color = "black",
              size = 3) +
    coord_radial(theta = "y",
                 expand = FALSE,
                 start = -pi / 2,
                 direction = -1L) +
    theme(axis.line = element_blank(),
          axis.ticks = element_blank(),
          axis.title = element_blank(),
          axis.text = element_blank())
print(p)
# ggsave(p, file = "tmp.pdf", width = 6, height = 6)
```

## Plot repeats as lollipop
```{r plotRawLollipop, fig.width = 7, fig.height = 3.5}
plt_d <- PrepRptPlotData()
for (sp in SpeciesNames(outgroup = TRUE)) {
    plt_d[, eval(sp) := get(sp) / 1e6]
}
levs <- plt_d[, repFamily[order(drum, decreasing = FALSE)]]
tmp_d <- plt_d[, .(repFamily, pilo)]
plt_d <- melt(plt_d,
              id.vars = "repFamily",
              variable.name = "SPECIES",
              value.name = "BP")
plt_d <- merge(plt_d, tmp_d, by = "repFamily")
plt_d[, RelativeOccupancy := log2((BP / nbases[SPECIES]) / (pilo / nbases["pilo"]))]
plt_d[, repFamily := factor(repFamily, levels = levs)]
plt_d[, SPECIES := factor(ConvertSpeciesNames(as.character(SPECIES),
                                              "short",
                                              "g_species"),
                          levels = ConvertSpeciesNames(c("roem", "drum", "cusp", "pilo"),
                                                       "short",
                                                       "g_species"))]

leg_lim <- c(1 / 3, 3)
leg_breaks <- c(1/2, 1, 2)
leg_labs <- sprintf("%0.1fX", leg_breaks)

p <- ggplot(plt_d[SPECIES != "P. pilosa"]) +
    geom_segment(aes(x = 0,
                     xend = BP,
                     y = repFamily,
                     yend = repFamily)) +
    geom_point(aes(x = BP,
                   y = repFamily,
                   fill = RelativeOccupancy),
               shape = 21,
               size = 3.7) +
    scale_fill_gradientn(colors = ryb_pal(100),
                         breaks = log2(leg_breaks),
                         labels = leg_labs,
                         limits = log2(leg_lim)) +
    facet_wrap(vars(SPECIES), nrow = 1) +
    xlab("Amount of sequence (Mb)") +
    ylab("Repeat family") +
    theme_bw() +
    guides(fill = guide_colorbar(title = "Sequence occupancy\nrelative to pilosa",
                                 direction = "horizontal")) +
    theme(legend.position = "top", panel.grid.major.y = element_blank(),
          strip.text = (fontface = element_text(face = "italic")))
print(p)

ggsave(p, file = "figure1.rpt_fam.pdf", width = 7, height = 3.5)
```

## Heatmap relative to pilosa

Plot overview of repeat family occupancy relative to pilosa

```{r plotRelativeHeat, fig.width = 3.5, fig.height = 4.5}
plt_d <- PrepRptPlotData()

# Convert to occupancy
for (sp in names(plt_d)[-1]) {
    plt_d[, eval(sp) := get(sp) / nbases[sp]]
}


title_str <- "Repeat family occupancy\nrelative to P. pilosa"
plt_m <- log2(as.matrix(plt_d[, .(cusp = cusp/pilo, drum = drum/pilo, roem = roem/pilo)]))

rownames(plt_m) <- plt_d[, repFamily]
rownames(plt_m)[which(rownames(plt_m) == "DNA")] <- "DNA/Unknown"
# annot_colors <- list("row")

maxval <- max(abs(range(plt_m)))
use_breaks <- seq(-maxval, maxval, length.out = 101)

leg_breaks <- c(1/3, 1/2, 1, 2, 3)
leg_labs <- sprintf("%0.1fX", leg_breaks)

#out_fn <- "te_fig.repFamily.indRepCount.heatmap.pdf"
pheatmap(plt_m, scale = "none", border_color = "#ffffff",
         color = colorRampPalette(rev(brewer.pal(n = 7, name = "RdYlBu")))(100),
         cluster_cols = TRUE,
         legend_breaks = log2(leg_breaks),
         legend_labels = leg_labs,
         breaks = use_breaks,
         # annotation_colors = annot_colors,
         treeheight_row = 25,
         treeheight_col = 10,
         main = title_str,
         # filename = out_fn,
         width = 3.5, height = 4.5)
```

# P-values

## Gene vs repeat overlaps

This was computed as part of the snakemake pipeline using 
1000 randomly shuffled gene intervals

```{r GeneRepeatOverlap}
pval <- fread("results/annotations/genes.rpt_overlap.pval.txt",
              header = FALSE)$V1
cat(str_interp("p = ${pval} (1000 replicates)\n"))
```

## Compute SV-intron/exon/intergenic location overlap

We make the simplifying assumption that SVs are points and so 
overlap is determined exclusively by the start position. So in 
effect we ignore cases where a deletion might extend into a 
feature interval.

### Load data

```{r LoadSvOverlaps}
sv_overlaps <- list()
for (feat in c("exons", "introns", "genes")) {
    fn <- str_interp("results/sv_overlap/${feat}.overlaps.bed")
    sv_overlaps[[feat]] <- fread(
        fn,
        header = FALSE,
        select = c(1, 3:6),
        col.names = c("CHROM", "POS", "ID", "SVLEN", "SVTYPE")
    )
    sv_overlaps[[feat]][, FEATURE := feat]
}

cnames <- names(sv_overlaps$genes)[-ncol(sv_overlaps$genes)]
sv_overlaps[["intergenic"]] <- sv_gts[!(ID %in% sv_overlaps$genes$ID),
                                      cnames,
                                      with = FALSE]
sv_overlaps[["intergenic"]][, FEATURE := "intergenic"]
sv_overlaps <- rbindlist(sv_overlaps)

uniq_bp <- c()
for (feat in c("exons", "introns", "genes")) {
    fn <- str_interp("results/annotations/${feat}.uniq_bp.txt")
    uniq_bp[feat] <- fread(fn, header = FALSE)$V1
}
fn <- "results/annotations/gaps.bed"
uniq_bp["gaps"] <- fread(fn, header = FALSE)[, sum(V3 - V2)]

fn <- "results/ref/phlox_flye.v1.0.FINAL.fasta.fai"
uniq_bp["total"] <- fread(fn, header = FALSE)[, sum(V2)] - uniq_bp["gaps"]

uniq_bp["intergenic"] <- uniq_bp["total"] - uniq_bp["genes"]
```

### Compute p-value
```{r SvOverlapPvalues}
sv_counts <- sv_overlaps[, table(FEATURE)]
overlap_tests <- list()
for (feat in names(sv_counts)) {
    overlap_tests[[feat]] <- list(FEATURE = feat)
    x <- sv_counts[feat]
    n <- nrow(sv_gts)
    overlap_tests[[feat]][["EXP"]] <- feat_prop <- uniq_bp[feat] / uniq_bp["total"]
    for(alt in c("less", "greater")){
        overlap_tests[[feat]][[alt]] <- binom.test(x, n, feat_prop, alt)$p.value
    }
    overlap_tests[[feat]][["OBS"]] <- x / n
    
}
overlap_tests <- rbindlist(overlap_tests)
```

### Plot SV occupancy per megabase
```{r PlotSvOverlap, fig.width=4, fig.height=4}
plt_d <- sv_overlaps[, .N, by = .(FEATURE, SVTYPE)]
setnames(plt_d, c("N", "SVTYPE"), c("SVs", "SV type"))

nms <- sv_overlaps[, unique(FEATURE)]
plt_d <- merge(plt_d, list(FEATURE=nms, BP=uniq_bp[nms]), by="FEATURE")

plt_d[, Y := SVs / (BP / 1e6)]

levs <- plt_d[, sum(Y), by = FEATURE][,
    FEATURE[order(V1, decreasing = TRUE)]]
plt_d[, FEATURE := factor(FEATURE, levels = levs)]
levs <- c("INS", "DEL", "INV", "DUP")
plt_d[, `SV type` := factor(`SV type`, levels = levs)]

annot_d <- overlap_tests[greater < 0.001]

annot_d <- merge(annot_d,
                 plt_d[, .(Y = sum(Y) + 10), by = FEATURE],
                 by = "FEATURE")

p <- ggbarplot(plt_d,
               x = "FEATURE",
               y = "Y",
               fill = "SV type",
               palette = "d3") +
    geom_text(aes(label = "***"), data = annot_d) +
    xlab("Genomic feature") +
    ylab("SVs per feature megabase")
print(p)
# ggsave(p, file = "tmp.pdf", width = 3, height = 4)
```

## Pi values in different partitions of the genome_plots

### Load Data
```{r LoadPiData}
pi_info <- list()
pi_info[["min_dp"]] <- 5

feat_levels <- c("exon", "intron", "near_gene", "intergenic")

tmp_d <- list()
for (feat in feat_levels) {
    cmd <- str_interp("tail -q -n+2 pi_stats/${feat}/merged.chr[1-8].ALL_REPS.DP_${pi_info$min_dp}.counts.tsv")
    x <- fread(cmd = cmd, header = FALSE, col.names = c("NUM", "DEN"))
    x[, REP := rep(seq(1, 5), 8)]
    tmp_d[[feat]] <- x[, .(PI = sum(NUM) / sum(DEN)), by = REP]
    tmp_d[[feat]][, FEATURE := feat]
}
pi_info[["data"]] <- rbindlist(tmp_d)
pi_info[["data"]][, FEATURE := factor(FEATURE, levels = feat_levels)]

tmp_d <- list()
for (feat in "exon") {
    fn <- str_interp("pi_stats/${feat}/merged.ALL_CHR.ALL_REPS.DP_${pi_info$min_dp}.bootstrap.pi.tsv")
    tmp_d[[feat]] <- fread(fn, header = TRUE)
    tmp_d[[feat]][, BOOT_I := .I]
    tmp_d[[feat]][, FEATURE := feat]
}
pi_info[["boot"]] <- rbindlist(tmp_d)
```

### P-value for exons

How reduced are exons relative to background? Used randomly 
shuffled exon intervals across genome to bootstrap a distribution

```{r PiExonPval}
pt_est <- pi_info$data[FEATURE == "exon", median(PI)]
nboot <- pi_info$boot[FEATURE == "exon", .N]
pval <- pi_info$boot[FEATURE == "exon", sum(PI < pt_est) / nboot]
pval <- str_interp(ifelse(pval == 0, "p < ${1 / nboot}", "p = ${pval}"))

cat("P-value for pi less than that observed in exons:\n", pval, "\n", sep="")
```

### Plot pi values
Note that near_genes is the 1kb up- and down-stream of a gene feature.
```{r plotPiSupplement, fig.width=4.5, fig.height=3}
plt_d <- pi_info$data[, QuantileOverReplicates(PI, prefixes = "PI"), by = FEATURE]
p <- ggplot(plt_d, aes(x = FEATURE, y = PI_med)) +
    geom_bar(aes(fill = FEATURE), stat = "identity") +
    geom_linerange(aes(ymin = PI_lwr, ymax = PI_upr), color = "black") +
    xlab("Feature") +
    ylab("Nucleotide diversity") +
    theme_pubr() +
    theme(legend.position = "none")

print(p)
print(plt_d)
```

# Save for debugging
```{r saveDebug}
save.image("results/figure1.Rdata")
#load("results/figure1.Rdata")
```
