---
title: "Figure 2"
date: "2025-03-13"
output: 
    html_document:
        toc: true
        toc_float: true
author: Felix L. Wu
---

# Setup

## Load packages

```{r loadPkg, message = FALSE}
source("../utils/genome_plots.R")

# Install packages
req_pkgs <- c("data.table",
              "stringr",
              "pracma",
              "ggpubr",
              "ggnewscale",
              "viridis",
              "patchwork",
              "extrafont",
              "digest",
              "poppr")
LoadPackages(req_pkgs)
LoadEEMSplotLib()

SetBaseTheme(base_family = "Arial", fontpath = "~/.local/share/fonts")
```

## Theme/font stylings
```{r themeSet, message = FALSE}
plist <- list()
# pc1_color_range <- c("#524C98", "#D1A4BB")
pc1_color_range <- c("#352E80", "#D699B7", "#FFDEEE")
eems_colors <- c("#994000", "#CC5800", "#FF8F33", "#FFAD66", "#FFCA99", "#FFE6CC",
                 "#DAE9FD", "#B5D3FB", "#90BDF9", "#6BA7F7", "#4691F4", "#217CF3")

# > dichromat::dichromat(colorRampPalette(pc1_color_range)(10))
#  [1] "#3F3F83" "#54548E" "#6C6C99" "#8686A4" "#A1A1AF" "#B4B4BB" "#C0C0C7"
#  [8] "#CDCDD3" "#D9D9DF" "#E6E6EB"
# > dichromat::dichromat(eems_colors)
#  [1] "#6B6B27" "#8A8A1F" "#B9B937" "#C9C966" "#DADA98" "#ECECCA" "#E3E3FB"
#  [8] "#CACAF9" "#B2B2F7" "#9A9AF5" "#8383F2" "#6E6EF1"
```


## Metadata
```{r loadMet}
met_fn <- "metadata.tsv"
met_d <- fread(met_fn, header = TRUE, sep = "\t", na.strings = c("NA", "#N/A"))
```

Figure out which cusp population is closest to each drum population

```{r drumCuspDist, fig.width = 4, fig.height = 3}
drum_pops <- met_d[species == "drum", unique(Population)]
tmp <- met_d[, .(lat = lat[1], long = long[1], species = species[1]), by = Population]
cusp_pops <- tmp[species == "cusp", Population]
setkey(tmp, Population)

nearest_cusp <- list()
for (i in drum_pops) {
    min_dist <- Inf
    drum_coords <- tmp[i, c(lat, long)]
    j <- tmp[cusp_pops, haversine(drum_coords, c(lat, long)), by = Population][, Population[which.min(V1)]]
    nearest_cusp[[i]] <- data.table(DRUM_POP = i,
                                    CUSP_POP = j,
                                    DIST = haversine(drum_coords, tmp[j, c(lat, long)]))
}
nearest_cusp <- rbindlist(nearest_cusp)
rm(tmp)

symp_dist_cutoff <- 30

p <- ggplot(nearest_cusp, aes(DIST)) +
    stat_ecdf(geom = "step") +
    geom_vline(xintercept = symp_dist_cutoff, color = "red") +
    xlab("Drum pop distance to nearest cusp")
print(p)

symp_drum <- nearest_cusp[DIST < symp_dist_cutoff, DRUM_POP]
```

## Function for getting a results directory from parameter settings
```{r resDir}
ResDirFromParams <- function(breadth, species, min_samp, mind) {
    require(stringr)
    str_interp("results/${breadth}/${species}/res_stacks/min_samples_per_pop~${min_samp}/min_samples_overall~${min_samp}/mind_${mind}")
}
```



# PCA

Both 3-species (all annuals) PCA and drummondii-only PCA.

## Load data

```{r prepPCA}
pca_info <- list()
for (nm in c("three", "drum")) {
    if (nm == "three") {
        res_dir <- ResDirFromParams("multi", "cusp_drum_roem", min_samp = 0.7, mind = 0.25)
    } else {
        res_dir <- ResDirFromParams("single", "drum", min_samp = 0.5, mind = 0.5)
    }
    pca_prefix <- str_interp("${res_dir}/pca_impute/populations.snps.lFilt.iFilt.pca.impute.missForest")
    pca_info[[nm]] <- LoadPCA(pca_prefix)
}
```

## Plot
```{r plotPCA, fig.width = 4.5, fig.height = 4.5}
plist[["pca"]] <- list()
for (nm in names(pca_info)) {
    plt_d <- merge(pca_info[[nm]]$evec, met_d, by.x = "IID", by.y = "Sample_ID")

    xyvar <- c("PC1", "PC2")

    axis_labs <- c()
    for (i in seq_along(xyvar)) {
        pct <- pca_info[[nm]]$prop_var[xyvar[i]] * 100
        axis_labs[i] <- str_interp("${xyvar[i]} ($[0.1f]{pct}%)")
    }

    if (nm == "three") {
        cvar <- "species"
        plt_d[, species := factor(ConvertSpeciesNames(species), levels = ConvertSpeciesNames(c("roem", "drum", "cusp")))]
    } else {
        cvar <- "PC1"
    }

    p <- ggscatter(plt_d, x = xyvar[1], y = xyvar[2], color = cvar, alpha = ifelse(nm == "three", 0.5, 0.75), font.family = "Arial") +
        xlab(axis_labs[1]) + ylab(axis_labs[2])
    if (nm == "three") {
        use_col <- Load4SpPal()
        names(use_col) <- ConvertSpeciesNames(names(use_col))
        p <- p +
            scale_color_manual(values = use_col) +
            guides(color = guide_legend(title = NULL))
    } else {
        p <- p + scale_color_gradient2(
            low = pc1_color_range[1],
            mid = pc1_color_range[2],
            high = pc1_color_range[3]
        )
    }

    plist[["pca"]][[nm]] <- p
    print(p)
}

```

# Admixture

## Load data
```{r prepAdmix}
res_dir <- ResDirFromParams("multi", "cusp_drum_roem", min_samp = 0.5, mind = 0.5)
admix_info <- list()
admix_info[["cv"]] <- fread(
    str_interp("${res_dir}/admix/populations.snps.lFilt.iFilt.admix.CV_report.txt"),
    sep = ":",
    header = FALSE,
    col.names = c("K", "CV_error")
)
admix_info[["cv"]][, K := as.integer(str_replace(str_replace(K, "CV error \\(K=", ""), "\\)", ""))]

samps <- fread(str_interp("${res_dir}/admix/populations.snps.lFilt.iFilt.admix.fam"), header = FALSE)$V2
samp_ord <- met_d[Sample_ID %in% samps,
                  .(Sample_ID,
                    species = factor(species, levels = c("roem", "drum", "cusp")),
                    long)]
setorder(samp_ord, species, long)
admix_info[["ord"]] <- samp_ord

for (k in 1:8) {
    nm <- as.character(k)
    tmp <- fread(
        str_interp("${res_dir}/admix/populations.snps.lFilt.iFilt.admix.${k}.Q"),
        header = FALSE
    )
    tmp[, IID := samps]

    tmp <- melt(tmp, id.vars = "IID", variable.name = "pop_infer")
    tmp[, pop_infer := str_replace(pop_infer, "V", "p")]
    tmp <- merge(
        tmp,
        met_d[, .(Sample_ID, species, Population, lat, long)],
        by.x = "IID",
        by.y = "Sample_ID"
    )

    tmp[, IID := factor(IID, levels = samp_ord$Sample_ID)]

    admix_info[[nm]] <- tmp
}

admix_info[["plot_k"]] <- "3"
```

## Plot

CV error for different k values

```{r plotAdmixError, fig.width = 4, fig.height = 3}
plt_d <- admix_info[["cv"]]
p <- ggplot(plt_d, aes(x = K, y = CV_error)) +
    geom_point(color = "dodgerblue") + geom_line(color = "dodgerblue") +
    ylab("Cross-validation error")

print(p)
plist[["admix_cv"]] <- p

ggsave(p, file = "Supplement.CV_Error.pdf",
       width = 5, height = 4,
       device = cairo_pdf, family = "Arial")
```

Actual admixture plot

```{r plotAdmix, fig.width = 8, fig.height = 2}
k <- admix_info[["plot_k"]]
plt_d <- admix_info[[k]]

# Identify admixed samples
tmp <- dcast(plt_d[, .(IID, pop_infer, value)], IID ~ pop_infer)
anc_co <- 0.9
admix_samps <- tmp[((p1 > anc_co) + (p2 > anc_co) + (p3 > anc_co)) != 1, IID]
admix_annot <- plt_d[IID %in% admix_samps, .(X = as.integer(IID), Y = cumsum(rev(value))[which.min(abs(cumsum(rev(value)) - 0.5))]), by = IID]


x_annot <- which(!duplicated(admix_info[["ord"]]$species))[-1] - 0.5
x_midpts <- (c(1, x_annot + 0.5) + c(x_annot + 0.5, length(levels(plt_d$IID)))) / 2

# Associate ancestry with species
if (as.integer(k) == 4) {
    tmp <- AdmixAnc2Species(plt_d)
    if (length(tmp$remain_pops) != 2) {
        stop("Failed to find a four-color palette for k=4")
    }
    admix_pal <- tmp$admix_pal
    admix_pal[tmp$remain_pops[1]] <- unname(Load4SpPal("2drum")["drumBlue"])
    admix_pal[tmp$remain_pops[2]] <- unname(Load4SpPal("2drum")["drumPink"])
} else if (as.integer(k) == 3) {
    tmp <- AdmixAnc2Species(plt_d)
    if (length(tmp$remain_pops) != 1) {
        stop("Failed to find a three-color palette for k=4")
    }
    admix_pal <- tmp$admix_pal
    admix_pal[tmp$remain_pops] <- unname(Load4SpPal()["drum"])
} else {
    admix_pal <- get_palette("d3", k = as.integer(k))
    names(admix_pal) <- str_c("p", 1:as.integer(k))
}

p <- ggplot(plt_d, aes(x = IID, y = value, fill = pop_infer)) +
    geom_bar(stat = "identity", width = 1) +
    geom_hline(yintercept = 0, color = "black") +
    scale_fill_manual(values = admix_pal) +
    ylab("Proportion inferred ancestry") +
    ggtitle(str_interp("k = ${k}")) +
    theme(
        axis.line.x = element_blank(),
        axis.title.x = element_blank(),
        axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        axis.title.y = element_text(size = 9),
        legend.title = element_text(size = 9),
        legend.position = "right"
    )
for (x in x_annot) {
    p <- p + annotate("path", x = rep(x, 2), y = c(0, 0.05), color = "black")
}

p <- p + annotate("point", x = admix_annot$X, y = admix_annot$Y, color = "white", shape = "*", size = 6)



for (i in seq_along(x_midpts)) {
    x <- x_midpts[i]
    sp <- admix_info[["ord"]][, levels(species)][i]
    p <- p + annotate("text", x = x, y = 0.05, label = ConvertSpeciesNames(sp), hjust = "center", color = "black", fontface = "italic", size = 3)
}
# ggsave(p, file = "tmp.pdf", width = 8, height = 2)

plist[["admix"]] <- p
print(p)
```

Supplementary admixture plots

```{r plotSuppAdmix, fig.width = 8, fig.height = 12}
k_values <- as.character(setdiff(2:8, admix_info$plot_k))
supp_plist <- list()

for (k in k_values) {
    plt_d <- admix_info[[k]]

    x_annot <- which(!duplicated(admix_info[["ord"]]$species))[-1] - 0.5
    x_midpts <- (c(1, x_annot + 0.5) + c(x_annot + 0.5, length(levels(plt_d$IID)))) / 2

    admix_pal <- get_palette("d3", k = as.integer(k))
    names(admix_pal) <- str_c("p", 1:as.integer(k))

    p <- ggplot(plt_d, aes(x = IID, y = value, fill = pop_infer)) +
        geom_bar(stat = "identity", width = 1) +
        geom_hline(yintercept = 0, color = "black") +
        scale_fill_manual(values = admix_pal) +
        ylab("Proportion inferred ancestry") +
        ggtitle(str_interp("k = ${k}")) +
        theme(
            axis.line.x = element_blank(),
            axis.title.x = element_blank(),
            axis.text.x = element_blank(),
            axis.ticks.x = element_blank(),
            axis.title.y = element_text(size = 9),
            legend.title = element_text(size = 9),
            legend.position = "right"
        )
    for (x in x_annot) {
        p <- p + annotate("path", x = rep(x, 2), y = c(0, 0.05), color = "black")
    }

    for (i in seq_along(x_midpts)) {
        x <- x_midpts[i]
        sp <- admix_info[["ord"]][, levels(species)][i]
        p <- p + annotate("text", x = x, y = 0.05, label = ConvertSpeciesNames(sp), hjust = "center", color = "black", fontface = "italic", size = 3)
    }

    supp_plist[[k]] <- p + theme(legend.position = "none")
}

p <- supp_plist[["2"]] /
    supp_plist[["4"]] /
    supp_plist[["5"]] /
    supp_plist[["6"]] /
    supp_plist[["7"]] /
    supp_plist[["8"]]


ggsave(p, file = "Supplement.Admix.pdf",
       width = 8, height = 12,
       device = cairo_pdf, family = "Arial")
print(p)
```


# EEMS

## Load data

```{r prepEEMS}
# Get EEMS results paths
eems_info <- list()
for (sp in SpeciesNames("short")) {
    res_dir <- ResDirFromParams("multi", "cusp_drum_roem", min_samp = 0.7, mind = 0.25)
    eems_info[[sp]] <- str_interp("${res_dir}/eems/${sp}/populations.snps.lFilt.iFilt.eems_out")
}

eems_info[["longlat"]]     <- FALSE
eems_info[["add_grid"]]    <- FALSE
eems_info[["add_demes"]]   <- FALSE
eems_info[["eems_colors"]] <- eems_colors
```


## Plot

```{r plotEEMS, warning = FALSE, message = FALSE, fig.width = 10, fig.height = 10}
AnnotTexasBorder <- function(p, bounding_box = TRUE, bounding_box_color = "black", thirds = TRUE, thirds_labs = FALSE) {
    require(ggplot2)
    require(data.table)
    L2 <- LONG <- LAT <- NULL # for linting
    texas <- LoadTexasBorder() # nolint: object_usage_linter.
    edplat <- LoadEdwardsPlateau() # nolint: object_usage_linter.
    p <- p + annotate(
        "path",
        x = texas[L2 == 1, LONG],
        y = texas[L2 == 1, LAT],
        color = "black"
    ) + annotate(
        "path",
        x = edplat[L2 == 1, LONG],
        y = edplat[L2 == 1, LAT],
        color = "brown",
        linetype = "dotted"
    )

    cur_lim <- GetPlotLimits(p) # nolint: object_usage_linter.

    if (thirds) {
        thirds_color <- "#999999"
        intercepts <- t(apply(cur_lim, 1, function(a) round(seq(a[1], a[2], length.out = 4)[2:3], 1)))
        cur_dim <- apply(cur_lim, 1, diff)
        for (i in 1:2) {
            p <- p + geom_vline(xintercept = intercepts["x", i], color = thirds_color, linetype = 2, linewidth = 0.7)
            p <- p + geom_hline(yintercept = intercepts["y", i], color = thirds_color, linetype = 2, linewidth = 0.7)

            if (thirds_labs) {
                p <- p + annotate(geom = "text", label = sprintf("'%0.1f'*degree", intercepts["x", i]),
                                  x = intercepts["x", i] - cur_dim["x"] * 0.01, y = cur_lim["y", 1] + cur_dim["y"] * 0.02,
                                  hjust = "right", vjust = "bottom",
                                  color = thirds_color, parse = TRUE, size = 4)

                p <- p + annotate(geom = "text", label = sprintf("'%0.1f'*degree", intercepts["y", i]),
                                  x = cur_lim["x", 1] + cur_dim["x"] * 0.018, y = intercepts["y", i] + cur_dim["y"] * 0.02,
                                  hjust = "left", vjust = "bottom",
                                  color = thirds_color, parse = TRUE, size = 4)
            }
        }
    }

    if (bounding_box) {
        p <- p + coord_cartesian(xlim = cur_lim["x", ], ylim = cur_lim["y", ], expand = FALSE) +
            theme(panel.border = element_rect(color = bounding_box_color, fill = NA, linewidth = 4))
    }

    return(p)
}

AnnotPopulations <- function(p, cur_samps, met_d, fill_data, aggr_func = "mean", color = "black") {
    require(data.table)
    require(ggnewscale)
    Sample_ID <- Population <- lat <- long <- V1 <- size <- . <- NULL # for linting # nolint: object_name_linter.

    annot_d <- met_d[Sample_ID %in% cur_samps, .(Sample_ID, Population, lat, long)]

    # Figure out which pops are the "same"
    tmp <- annot_d[, .(lat = round(lat[1], 4), long = round(long[1], 4)), by = Population]
    collapse_pops <- strsplit(tmp[, str_c(Population, collapse = ","), by = .(lat, long)][str_detect(V1, ","), V1], ",")
    for (i in seq_along(collapse_pops)) {
        nm <- str_c(collapse_pops[[i]], collapse = "/")
        annot_d[Population %in% collapse_pops[[i]], Population := nm]
    }

    if (!is.null(fill_data)) {
        annot_d <- merge(annot_d, fill_data, by = "Sample_ID")
        fill_var <- names(fill_data)[names(fill_data) != "Sample_ID"]

        lambda <- eval(parse(text = aggr_func))
        tmp <- annot_d[, lambda(get(fill_var)), by = Population]$V1
        annot_d <- annot_d[, .(size = .N, lat = lat[1], long = long[1]), by = Population]
        annot_d[, eval(fill_var) := tmp]

        p <- p + new_scale_fill() + geom_point(aes(x = long, y = lat, size = size, fill = .data[[fill_var]]), data = annot_d, shape = 21, color = color)
    } else {
        annot_d <- annot_d[, .(size = .N, lat = lat[1], long = long[1]), by = Population]
        p <- p + geom_point(aes(x = long, y = lat, size = size), data = annot_d, fill = "black", shape = 21, color = color, alpha = 0.5)
    }
    p <- p + guides(size = "none")

    return(p)
}

plist[["eems"]] <- list()
for (sp in SpeciesNames("short")) {
    # Save plot list object for later retrieval since this can take a while to run.
    tmp_args <- eems_info[names(eems_info)[!(names(eems_info) %in% SpeciesNames("short"))]]
    tmp_args[["mcmcpath"]] <- eems_info[[sp]]
    dig_str <- digest(object = tmp_args)
    rdat_fn <- str_interp("${eems_info[[sp]]}/${dig_str}.Rdata")
    if (file.exists(rdat_fn)) {
        load(rdat_fn)
    } else {
        p <- do.call(make_eems_plots, tmp_args)
        save(p, file = rdat_fn)
    }

    if (sp == "drum"){
        tmp <- as.data.table(p$mrates01$data)
        ed_plat_pops <- rbind(c(-98.54003, 30.29176), c(-98.841583, 30.272333), c(-98.110483, 30.5497))
        logm_vals <- c()
        for (i in seq_len(nrow(ed_plat_pops))) {
            xcoord <- targvals[i, 1]
            ycoord <- targvals[i, 2]
            logm_vals <- c(logm_vals, tmp[abs(ycoord - y) < 0.002][which.min(abs(xcoord - x)), z])
        }
        cat("Southeastern Edwards Plateau population log(m):\n")
        print(logm_vals)
    }

    for (nm in names(p)) {
        if (str_detect(nm, "rates")) {
            p[[nm]] <- AnnotTexasBorder(p[[nm]], bounding_box_color = Load4SpPal()[sp], thirds_labs = (sp == "drum"))
            cur_samps <- fread(str_interp("${eems_info[[sp]]}/../populations.snps.lFilt.iFilt.eems.fam"), header = FALSE)$V2
            if (sp == "drum" && nm == "mrates01") {
                fill_data <- pca_info$drum$evec[, .(Sample_ID = IID, PC1)]
                p[[nm]] <- AnnotPopulations(p[[nm]], cur_samps, met_d, fill_data = fill_data) +
                    scale_fill_gradient2(name = "Mean PC1",
                                         limits = range(fill_data$PC1),
                                         low = pc1_color_range[1],
                                         mid = pc1_color_range[2],
                                         high = pc1_color_range[3])
            } else if (sp == "drum" && nm == "qrates01") {
                # Change default color scale from blue-orange to cividis
                cividis_endcolors <- cividis(2)
                p[[nm]] <- p[[nm]] + scale_fill_gradient2(name = "log(q)",
                                                          low = cividis_endcolors[1],
                                                          mid = "white",
                                                          high = cividis_endcolors[2],
                                                          limits = c(-0.1, 0.1))

                fill_data <- pca_info$drum$evec[, .(Symp = ifelse(FID %in% symp_drum, "sympatry", "allopatry")), by = IID]
                setnames(fill_data, "IID", "Sample_ID")

                use_colors <- Load4SpPal("2drum")[c("drumBlue", "drumPink")]
                names(use_colors) <- c("allopatry", "sympatry")
                p[[nm]] <- AnnotPopulations(p[[nm]],
                                            cur_samps, met_d,
                                            fill_data = fill_data,
                                            aggr_func = "function(x) x[1]") +
                    scale_fill_manual(name = "", values = use_colors)
            } else {
                p[[nm]] <- AnnotPopulations(p[[nm]], cur_samps, met_d, fill_data = NULL)
            }
        }
    }
    plist[["eems"]][[sp]] <- p
    p_m01 <- p[["mrates01"]] + ggtitle(sp)
    print((p_m01 | p[["mrates02"]]) / (p[["qrates01"]] | p[["qrates02"]]))
}
```

# Isolation by distance

## Load data
```{r loadFst}
popwise_fst <- list()
for (sp in SpeciesNames("short")) {
    res_dir <- ResDirFromParams("single", sp, min_samp = 0.5, mind = 0.5)
    fn <- str_interp("${res_dir}/fis/populations.all.lFilt.iFilt_fis.txt")
    popwise_fst[[sp]] <- fread(fn, header = TRUE, sep = "\t")
    popwise_fst[[sp]][, SPECIES := sp]
    setnames(popwise_fst[[sp]], letters[1:3], LETTERS[1:3])

    # Remove single individual populations
    fn <- str_interp("${res_dir}/fis/populations.all.lFilt.iFilt_fis.txt.pop")
    pop_d <- fread(fn, header = FALSE, col.names = c("IID", "FID"))
    rm_pops <- pop_d[, length(IID) == 1, by = FID][(V1), FID]
    popwise_fst[[sp]] <- popwise_fst[[sp]][!(pop1 %in% rm_pops | pop2 %in% rm_pops)]
}
popwise_fst <- rbindlist(popwise_fst)
all_fst <- popwise_fst[!is.na(A)][!is.na(B)][!is.na(C)][, .(FST = sum(A) / sum(A + B + C)), by = .(SPECIES, pop1, pop2)]
all_fis <- popwise_fst[!is.na(B)][!is.na(C)][, .(FIS = 1 - sum(C) / sum(B + C)), by = .(SPECIES, pop1, pop2)]

# We need pairwise distances between drum populations
pop_loc <- met_d[species %in% SpeciesNames("short"), .(lat = lat[1], long = long[1]), by = Population]
all_fst <- merge(all_fst, pop_loc, by.x = "pop1", by.y = "Population", all.x = TRUE)
setnames(all_fst, c("lat", "long"), c("lat1", "long1"))

all_fst <- merge(all_fst, pop_loc, by.x = "pop2", by.y = "Population", all.x = TRUE)
setnames(all_fst, c("lat", "long"), c("lat2", "long2"))

# Compute distance
all_fst[, DIST := haversine(c(lat1, long1), c(lat2, long2)), by = .(pop1, pop2)]
all_fst[, max_long := max(long1, long2), by = .(pop1, pop2)]

pop_four <- fread("drum.pop_categories.4grp.tsv", header = TRUE)
all_fst <- merge(all_fst, pop_four, by.x = "pop1", by.y = "Population", all.x = TRUE)
all_fst[is.na(Four_groups), Four_groups := "none"]
setnames(all_fst, "Four_groups", "Four_groups1")
all_fst <- merge(all_fst, pop_four, by.x = "pop2", by.y = "Population", all.x = TRUE)
all_fst[is.na(Four_groups), Four_groups := "none"]
setnames(all_fst, "Four_groups", "Four_groups2")
```

## Plot

As individual figures

```{r plotIsoByDist, fig.width = 5, fig.height = 4}
# Separate plots
plist[["ibd"]] <- list()
for (sp in SpeciesNames("short")) {
    p <- ggscatter(
        all_fst[SPECIES == sp],
        x = "DIST",
        y = "FST",
        add = "reg.line",
        conf.int = TRUE,
        color = "max_long",
        cor.coef = TRUE,
        alpha = 0.5,
        xlab = "Population distance (km)",
        ylab = "Fst",
        font.family = "Arial"
    )
    p <- p + scale_color_viridis(option = "plasma")

    plist[["ibd"]][[sp]] <- p
    print(p)
}
```

and as a faceted plot. We're also going to run a regression
for only FST between sympatric drummondii and non-sympatric drummondii.

```{r plotIsobyDistFact, fig.width = 12, height = 4}
include_allosymp <- TRUE
# Faceted plot
# get ordering correct...
all_fst[, SPECIES := factor(SPECIES, levels = c("roem", "drum", "cusp"))]

all_fst_lm <- list()
annot_d <- list()
label_d <- list()
for (sp in levels(all_fst$SPECIES)) {
    all_fst_lm[[sp]] <- RunOLS(all_fst[SPECIES == sp], "DIST", "FST")
    if (sp == "drum") {
        all_fst_lm[["allo_symp"]] <- RunOLS(all_fst[((Four_groups1 == "WestRed") + (Four_groups2 == "WestRed")) == 1], "DIST", "FST")
    }

    if (sp == "cusp") {
        all_fst_lm[["cusp_no_668"]] <- RunOLS(all_fst[SPECIES == sp][(pop1 != "668") & (pop2 != "668")], "DIST", "FST")
    }

    annot_d[[sp]] <- all_fst_lm[[sp]][["fitdat"]]
    annot_d[[sp]][, SPECIES := sp]
    annot_d[[sp]][, LABEL   := sp]

    lab_pos <- all_fst[SPECIES == sp, c(min(DIST), max(FST))]
    stat_str <- FormatR2PvalLab(unlist(all_fst_lm[[sp]][c("rsq", "pval")]))
    label_d[[sp]] <- data.table(x = lab_pos[1], y = lab_pos[2], lab = stat_str, SPECIES = sp)
}
label_d <- rbindlist(label_d)
label_d[, SPECIES := factor(SPECIES, levels = c("roem", "drum", "cusp"))]

if (include_allosymp) {
    annot_d[[4]] <- all_fst_lm[["allo_symp"]][["fitdat"]]
    annot_d[[4]][, SPECIES := "drum"]
    annot_d[[4]][, LABEL   := "allo vs symp"]

    use_colors <- Load4SpPal("2drum")
    col_brks <- c("roem", "drum", "allo vs symp", "cusp")
    col_labs <- c(ConvertSpeciesNames(col_brks[1:2]), "P. drummondii: allopatric vs. sympatric", ConvertSpeciesNames(col_brks[4]))
    names(use_colors)[names(use_colors) == "drumBlue"] <- "drum"
    names(use_colors)[names(use_colors) == "drumPink"] <- "allo vs symp"

    annot_d <- rbindlist(annot_d)
    annot_d[, SPECIES := factor(SPECIES, levels = col_brks[col_brks != "allo vs symp"])]
    annot_d[, LABEL   := factor(LABEL,   levels = col_brks)]
} else {
    use_colors <- Load4SpPal()
    col_brks <- c("roem", "drum", "cusp")
    col_labs <- ConvertSpeciesNames(col_brks)

    annot_d <- rbindlist(annot_d)
    annot_d[, SPECIES := factor(SPECIES, levels = col_brks)]
    annot_d[, LABEL   := factor(LABEL,   levels = col_brks)]
}

p <- ggplot() +
    geom_point(aes(x = DIST, y = FST, color = SPECIES), data = all_fst, alpha = 0.5) +
    geom_ribbon(aes(x = DIST, ymin = LWR, ymax = UPR, fill = LABEL), data = annot_d, alpha = 0.3, color = NA) +
    geom_line(aes(x = DIST, y = FST, color = LABEL), data = annot_d, linewidth = 1) +
    geom_hline(yintercept = 0.15, color = "darkgrey", linetype = "dashed") +
    scale_fill_manual(values = use_colors, guide = "none") +
    scale_color_manual(values = use_colors, name = "Regression", breaks = col_brks, labels = col_labs) +
    geom_text(aes(x = x, y = y, label = lab), data = label_d, parse = TRUE, hjust = 0, family = "Arial")#, vjust = 1)

if (include_allosymp) {
    tmp_d <- label_d[SPECIES == "drum"]
    tmp_d[, y := 0.9 * y]
    tmp_d[, lab := FormatR2PvalLab(unlist(all_fst_lm[["allo_symp"]][c("rsq", "pval")]))]
    p <- p + geom_text(aes(x = x, y = y, label = lab),
                       data = tmp_d,
                       parse = TRUE,
                       color = use_colors["allo vs symp"],
                       hjust = 0,
                       family = "Arial")
}

p <- p +
    facet_wrap(vars(SPECIES), nrow = 1, scales = "free") +
    xlab("Distance between population pairs (km)") +
    # xlab("Population pair distance (km)") +
    ylab("Fst") +
    theme(strip.background = element_blank(), strip.text.x = element_blank())

plist[["ibd"]][["all"]] <- p
print(p)
```

Finally, plot just drummondii data but separated by whether a sympatric population
is involved

```{r plotIsoByDistSymp, fig.width = 9, fig.height = 3}
plt_d <- all_fst[SPECIES == "drum"]
plt_d[, SYMP := "none"]
plt_d[pop1 %in% symp_drum | pop2 %in% symp_drum, SYMP := "one"]
plt_d[pop1 %in% symp_drum & pop2 %in% symp_drum, SYMP := "both"]
plt_d[, SYMP := factor(SYMP, levels = c("none", "one", "both"))]

pal <- Load4SpPal("2drum")
use_colors <- c("none" = unname(pal["drumBlue"]), "one" = "#D1A4BB", "both" = unname(pal["drumPink"]))

p <- ggscatter(
    plt_d,
    x = "DIST",
    y = "FST",
    color = "SYMP",
    cor.coef = TRUE,
    alpha = 0.7,
    xlab = "Population distance (km)",
    ylab = "Fst",
    font.family = "Arial"
)

p <- p +
    geom_smooth(method = "lm") +
    scale_color_manual(values = use_colors) +
    # scale_color_manual(name = "Sympatric pops", values = use_colors) +
    facet_wrap(vars(SYMP), nrow = 1) +
    guides(color = "none")
print(p)
plist[["ibd"]][["symp"]] <- p
```
## P-value testing

Does the "in sympatry" variable for *P. drummondii* significantly affect our regression model?
```{r pvalIsoByDistSymp}
set.seed(18318)
reg_d <- all_fst[SPECIES == "drum"]
nboot <- 500
nsubsamp <- reg_d[, sum(((Four_groups1 == "WestRed") + (Four_groups2 == "WestRed")) == 1)]
boot_coef <- replicate(RunOLS(reg_d[sample.int(.N, size = nsubsamp, replace = TRUE)], "DIST", "FST")$model$coefficients, n = nboot)
pval <- c(sum(boot_coef[1,] < all_fst_lm$allo_symp$model$coefficients[1]),
          sum(boot_coef[2,] < all_fst_lm$allo_symp$model$coefficients[2])) / nboot
cat(str_interp("Bootstrap one-sided pval of intercept less than observed: ${pval[1]}\n"))
cat(str_interp("Bootstrap one-sided pval of slope less than observed: ${pval[2]}\n"))
```

# AMOVA

Excoffier 1992 (https://pmc.ncbi.nlm.nih.gov/articles/PMC1205020/pdf/ge1312479.pdf)

Using Population as the grouping variable. For drummondii, also test Allo vs Symp

## Run
```{r runAmova}
RunAmova <- function(prefix, grp_dat, id_col = "Sample_ID", force = FALSE) {
    require(digest)
    require(stringr)
    require(data.table)
    require(adegenet)
    require(poppr)
    ID <- IID <- NULL # for linter

    dig_str <- str_c( # nolint: object_usage_linter.
        system(str_interp("md5sum ${prefix}.tped | cut -f 1 -d ' '"), intern = TRUE),
        digest(object = grp_dat),
        sep = "."
    )
    dname <- dirname(prefix) # nolint: object_usage_linter.
    rdat_fn <- str_interp("${dname}/${dig_str}.Rdata")

    if (file.exists(rdat_fn) && !force) {
        cat(str_interp("Loading data from ${rdat_fn}\n"))
        load(rdat_fn)
        return(amova_res)
    }

    fam <- fread(str_interp("${prefix}.tfam"), header = FALSE, select = 1:2)
    names(fam) <- c("FID", "IID")
    fam <- fam[IID %in% grp_dat[[id_col]]]

    tmp_prefix <- str_c(tempdir(), "/file") # nolint: object_usage_linter.
    fwrite(fam, file = str_interp("${tmp_prefix}.keep"), row.names = FALSE, col.names = FALSE, sep = "\t", quote = FALSE)

    system(str_interp("${plink_path} --tfile ${prefix} --allow-extra-chr --keep ${tmp_prefix}.keep --recode --out ${tmp_prefix}"))
    system(str_interp("${plink_path} --tfile ${prefix} --allow-extra-chr --keep ${tmp_prefix}.keep --recode 12 A --out ${tmp_prefix}"))

    genlight_obj <- adegenet::read.PLINK(str_interp("${tmp_prefix}.raw"), map.file = str_interp("${tmp_prefix}.map"))

    setnames(grp_dat, id_col, "ID")
    setkey(grp_dat, ID)

    newstrata <- grp_dat[adegenet::indNames(genlight_obj), ]
    newstrata[, ID := NULL]

    adegenet::strata(genlight_obj) <- newstrata
    form_str <- str_c("~", str_c(colnames(newstrata), collapse = "/"), sep = "")
    use_form <- as.formula(form_str)
    cat(str_interp("Using formula ${form_str}\n"))
    amova_res <- poppr.amova(genlight_obj, hier = use_form, within = TRUE, method = "pegas", nperm = 500)
    save(amova_res, file = rdat_fn)
    return(amova_res)
}

amova_info <- list()
for (sp in SpeciesNames()) {
    res_dir <- ResDirFromParams("multi", "cusp_drum_roem", min_samp = 0.7, mind = 0.25)
    tfile_prefix <- str_interp("${res_dir}/pca_impute/populations.snps.lFilt.iFilt.pca.impute.missForest")

    grp_dat <- met_d[species == sp, .(Sample_ID, Population)]
    amova_info[[sp]] <- RunAmova(tfile_prefix, grp_dat)#, force = TRUE)

    if (sp == "drum") {
        grp_dat <- met_d[species == sp, .(Sample_ID, Population)]
        grp_dat <- merge(grp_dat, pop_four, by = "Population", all.x = TRUE)
        grp_dat[, Region := "Allopatry"]
        grp_dat[Four_groups == "WestRed", Region := "Sympatry"]
        grp_dat[, Four_groups := NULL]
        setcolorder(grp_dat, c("Region", "Population", "Sample_ID"))
        amova_info[["allo_symp"]] <- RunAmova(tfile_prefix, grp_dat)#, force = TRUE)
    }
}
```

## Report results
```{r rptAmova}
for (nm in names(amova_info)) {
    x <- amova_info[[nm]]$varcomp
    x <- cbind(x, "PctVar" = x[, 1] / sum(x[, 1]) * 100)
    cat("\n", nm, "\n", sep = "")
    print(x)
}
```

## Plot

Percent explained ...
```{r plotAmovaPct, fig.width = 6, fig.height = 3}
plist[["amova"]] <- list()

lvls <- c("Between pop", "Within pop, between ind", "Within ind")
use_colors <- colorRampPalette(c("#333333", "#FFFFFF"))(3)
names(use_colors) <- lvls

plt_d <- list()
for (sp in SpeciesNames()) {
    y <- amova_info[[sp]]$varcomp[, 1]
    y <- y / sum(y)
    plt_d[[sp]] <- data.table(SPECIES = sp, PART = lvls, Y = y)
}

plt_d <- rbindlist(plt_d)
plt_d[, PART := factor(PART, levels = lvls)]
plt_d[, SPECIES := factor(ConvertSpeciesNames(SPECIES),
                          levels = ConvertSpeciesNames(c("cusp", "drum", "roem")))]

p <- ggbarplot(plt_d, x = "SPECIES", y = "Y", fill = "PART") +
    scale_fill_manual(name = "Partition", values = use_colors) +
    xlab("") + ylab("Proportion genetic variance explained") +
    coord_flip() +
    theme(axis.text.y = element_text(face = "italic"))
print(p)

plist[["amova"]][["norm"]] <- p
```

Raw ...

```{r plotAmovaRaw, fig.width = 6, fig.height = 3}
use_colors <- Load4SpPal()
names(use_colors) <- ConvertSpeciesNames(names(use_colors))

plt_d <- list()
for (sp in SpeciesNames()) {
    y <- sum(amova_info[[sp]]$varcomp[, 1])
    plt_d[[sp]] <- data.table(SPECIES = sp, Y = y)
}

plt_d <- rbindlist(plt_d)
plt_d[, SPECIES := factor(ConvertSpeciesNames(SPECIES),
                          levels = ConvertSpeciesNames(c("cusp", "drum", "roem")))]

p <- ggbarplot(plt_d, x = "SPECIES", y = "Y", fill = "SPECIES", color = NA) +
    scale_fill_manual(name = "Partition", values = use_colors) +
    xlab("") + ylab(expression(sigma^2)) +
    coord_flip() +
    theme(axis.text.y = element_text(face = "italic"),
          legend.position = "none")
print(p)

plist[["amova"]][["raw"]] <- p
```

# FIS

## Plot

(Data already loaded) when loading FST data

```{r plotFIS, fig.width = 5, fig.height = 6}
all_fis[, SPECIES := factor(SPECIES, levels = c("roem", "drum", "cusp"))]

addparams <- list(fill = NA, color = "black", width = 0.1)
p <- ggviolin(all_fis,
              x = "SPECIES", y = "FIS",
              color = "black", fill = "SPECIES",
              add = "median_iqr", add.params = addparams, error.plot = "crossbar") +
    scale_fill_manual(values = Load4SpPal()) #+
# ggsave(p, file = "tmp.pdf")
print(p)
```

# Inbreeding coefficients

## Load data
```{r loadHet}
het_info <- list()
for (sp in SpeciesNames("short")) {
    res_dir <- ResDirFromParams("single", sp, min_samp = 0.5, mind = 0.5)
    fn <- str_interp("${res_dir}/vcfhet/populations.snps.lFilt.iFilt.het")
    het_info[[sp]] <- fread(fn, header = TRUE)
    het_info[[sp]][, SPECIES := sp]
}
het_info <- rbindlist(het_info)
setnames(het_info, c("INDV", "O(HOM)", "E(HOM)", "F"), c("IID", "O_HOM", "E_HOM", "FSTAT"))

het_info[, SPECIES := factor(SPECIES, levels = c("roem", "drum", "cusp"))]
```

When plotting filter out possible hybrids

```{r plotHet, fig.width = 3.5, fig.height = 4}
# First remove the samples that appear to be admixed (from admixture code above)
rm_samps <- admix_samps
plt_d <- het_info[!(IID %in% rm_samps)]

detect_outliers <- FALSE
if (detect_outliers) {
    p <- ggboxplot(plt_d, x = "SPECIES", y = "FSTAT", fill = "SPECIES", alpha = 0.7)
} else {
    plt_d <- plt_d[, as.list(quantile(FSTAT, (0:4) / 4)), by = SPECIES]
    setnames(plt_d,
        str_c((0:4) /  4 * 100, "%"),
        c("MIN", "Q1", "MED", "Q3", "MAX")
    )
    p <- ggplot(plt_d, aes(x = SPECIES, ymin = MIN, lower = Q1, middle = MED, upper = Q3, ymax = MAX, fill = SPECIES)) +
        geom_boxplot(stat = "identity", alpha = 0.7)
}
p <- p + scale_fill_manual(name = "species", values = Load4SpPal()) +
    ylab("Inbreeding coefficient (F)") +
    theme(legend.position = "none",
          axis.title.x = element_blank())
plist[["het"]] <- p
print(p)
# ggsave(p, file = "tmp.pdf", width = 3.5, height = 4)
```


# Pi, dxy

## Load data
```{r loadPixy}
res_dir <- ResDirFromParams("multi", "cusp_drum_roem", min_samp = 0.1, mind = 0.9)
pixy_info <- list()

fn <- str_interp("${res_dir}/pixy/populations.all.lFilt.iFilt_pi.txt")
pixy_info[["pi_w"]] <- fread(fn, header = TRUE)
setnames(pixy_info[["pi_w"]], "pop", "SPECIES")
pixy_info[["pi_w"]][, SPECIES := factor(SPECIES, levels = c("roem", "drum", "cusp"))]

pixy_info[["pi_a"]] <- pixy_info[["pi_w"]][!is.na(count_diffs), BootstrapWtAvg(count_diffs, count_comparisons), by = .(SPECIES)]
print(pixy_info[["pi_a"]])


fn <- str_interp("${res_dir}/pixy/populations.all.lFilt.iFilt_dxy.txt")
pixy_info[["dxy_w"]] <- fread(fn, header = TRUE)

pixy_info[["dxy_a"]] <- pixy_info[["dxy_w"]][!is.na(count_diffs), BootstrapWtAvg(count_diffs, count_comparisons), by = .(pop1, pop2)]
print(pixy_info[["dxy_a"]])
```

# Attempts at full figure

## Version 1
... Removed ...

## Version 2
```{r plotFullFig2, fig.width = 14, fig.height = 14}
# Row 1
pPcaAll <- plist$pca$three + theme(legend.position = c(0.8, 0.8), # nolint: object_name_linter.
                                   legend.direction = "vertical",
                                   legend.text = element_text(face = "italic"),
                                   legend.background = element_rect(fill = "white", color = "black"))
pAdmix  <- plist$admix + theme(title = element_blank(), legend.position = "none") # nolint: object_name_linter.
pAmovaRaw  <- plist$amova$raw + theme(panel.background = element_blank()) # nolint: object_name_linter.
pAmovaNorm  <- plist$amova$norm + theme(axis.title.y = element_blank(), # nolint: object_name_linter.
                                        axis.text.y = element_blank(),
                                        panel.background = element_blank(),
                                        legend.position = "right")

# Row 2
pPcaDrum <- plist$pca$drum + theme(legend.position = c(0.5, 0.9), legend.direction = "horizontal", legend.text = element_text(size = 8)) # nolint: object_name_linter.
pEemsMigDrum <- plist$eems$drum$mrates01 + # nolint: object_name_linter.
    annotate("text", x = -Inf, y = Inf, hjust = -0.25, vjust = 1.7, label = ConvertSpeciesNames("drum"),
             color = Load4SpPal()["drum"], fontface = "italic", size = 7) +
    theme(legend.position = c(0.88, 0.8), legend.direction = "vertical") + guides(fill = "none")
pEemsMigCusp <- plist$eems$cusp$mrates01 + # nolint: object_name_linter.
    annotate("text", x = -Inf, y = Inf, hjust = -0.25, vjust = 1.7, label = ConvertSpeciesNames("cusp"),
             color = Load4SpPal()["cusp"], fontface = "italic", size = 7) +
    theme(legend.position = "none")
pEemsMigRoem <- plist$eems$roem$mrates01 + # nolint: object_name_linter.
    annotate("text", x = -Inf, y = Inf, hjust = -0.25, vjust = 1.7, label = ConvertSpeciesNames("roem"),
             color = Load4SpPal(for_text = TRUE)["roem"], fontface = "italic", size = 7) +
    theme(legend.position = "none")

# Row 3
pIbdAll <- plist$ibd$all # nolint: object_name_linter.

row_h <- cumsum(c(30, 25, 30))
col_w <- cumsum(c(30, 54))
col2_splitpoint <- col_w[1] + (diff(col_w) * 2 / 3)

# area(top, left, bottom, right)
layout <- c(
    area(1,  1, row_h[1], col_w[1]),   area(1,  col_w[1] + 1, row_h[1] / 2, col_w[2]), # PCA, Structure
    area(row_h[1] / 2 + 1, col_w[1] + 1, row_h[1], round(col_w[1] + diff(col_w) / 4)), # AMOVA (raw)
    area(row_h[1] / 2 + 1, round(col_w[1] + diff(col_w) / 4) + 1, row_h[1], col_w[2]), # AMOVA (norm)
    area(row_h[1] + 1, 1, row_h[2], col_w[2]), # IBD (3 subfigs)
    area(row_h[2] + 1, 1, row_h[3], col_w[1]), area(row_h[2] + 1, col_w[1] + 1, row_h[3], col2_splitpoint), # PCA (drum), EEMS (drum)
    area(row_h[2] + 1, col2_splitpoint + 1, mean(row_h[2:3]), col_w[2]), area(mean(row_h[2:3]) + 1, col2_splitpoint + 1, row_h[3], col_w[2]) # EEMS (cusp), EEMS (roem)
)


p <- pPcaAll + pAdmix + free(pAmovaRaw, side = "l") + free(pAmovaNorm, side = "r") +
    pIbdAll +
    pPcaDrum + pEemsMigDrum + pEemsMigCusp + pEemsMigRoem +
    plot_layout(design = layout)

ggsave(p, file = "figure2.pdf", width = 14, height = 14, device = cairo_pdf, family = "Arial")
print(p)
```

To do:
- Add species labels to second PCA, EEMS plots
- Something to go along with Inbreeding Coefficient plot?

# Save for debugging
```{r save}
save.image("results/figure2.Rdata")
# load("results/figure2.Rdata")
```
