---
title: "Figure 4 Dstat"
date: "2025-05-12"
output:
    html_document:
        toc: true
        toc_float: true
author: Felix L. Wu
---

# Setup

## Load packages

```{r loadPkg, message = FALSE}
source("../utils/genome_plots.R")

# Install packages
req_pkgs <- c("data.table", "stringr", "ggpubr", "viridis", "patchwork", "extrafont", "RColorBrewer")
LoadPackages(req_pkgs)
SetBaseTheme(base_family = "Arial", fontpath = "../utils/fonts")
plist <- list()
```

## Sample data
```{r getSamples}
pop_d <- fread("sample_lists/foursp.pop", header = FALSE, col.names = c("ID_LONG", "POP"))
pop_d[, ID_MED := str_remove(ID_LONG, ".*_Hopkins_")]
pop_d[, ID_SHORT := str_remove(str_remove(ID_MED, "_.*"), "-.*")]
pop_d[, ID_SHORT := str_c(toupper(substr(POP, 1, 1)), ID_SHORT)]
setkey(pop_d, ID_LONG)
```

## Functions and settings
```{r funcs}
ShortenID <- function(x, newid = "ID_SHORT") {
    stopifnot(newid %in% names(pop_d))
    return(pop_d[x][[newid]])
}
```

```{r globals}
vartypes <- c("SNP", "SV")
```

# D-statistic

## Load data

```{r loadDstat}
dstat_info <- list()
dstat_info[["DP"]] <- 5
dstat_info[["SVLEN"]] <- 500
dstat_info[["wsize"]] <- "250kb"

dstat_info[["w"]] <- list()
for (i in 1:5) {
    v <- "SNP"
    for (chunk in 1:16) {
        cmd <- str_interp("grep -v '^#' results/snp_patt/merged.chunk${chunk}.mpileup.rep${i}.DP_${dstat_info$DP}.patt_counts.w_${dstat_info$wsize}.tsv")
        cat(cmd, "\n", sep = "")
        dstat_info[["w"]][[cmd]] <- fread(cmd = cmd, header = TRUE, sep = "\t")
        dstat_info[["w"]][[cmd]][, REP := i]
        dstat_info[["w"]][[cmd]][, VARTYPE := v]
        dstat_info[["w"]][[cmd]][, CHROM := as.character(CHROM)]
    }

    v <- "SV"
    cmd <- str_interp("grep -v '^#' results/svs_patt/svs.rep${i}.SVLEN_${dstat_info$SVLEN}.patt_counts.w_${dstat_info$wsize}.tsv")
    cat(cmd, "\n", sep = "")
    dstat_info[["w"]][[cmd]] <- fread(cmd = cmd, header = TRUE, sep = "\t")
    dstat_info[["w"]][[cmd]][, REP := i]
    dstat_info[["w"]][[cmd]][, VARTYPE := v]
    dstat_info[["w"]][[cmd]][, CHROM := str_remove(str_remove(CHROM, "_RagTag$"), "^chrlg")]
}

dstat_info[["w"]] <- rbindlist(dstat_info[["w"]])
dstat_info[["w"]][, CHROM := factor(CHROM, levels = as.character(1:7))]
dstat_info[["all"]] <- dstat_info[["w"]][,
                                         .(N_SNPS = sum(N_SNPS), BBAA = sum(BBAA), ABBA = sum(ABBA), BABA = sum(BABA)),
                                         by = .(ID1, ID2, ID3, REP, VARTYPE)]
```

Compute D

```{r calcDstatSnp}
dstat_info$w[, DSTAT := sum(ABBA - BABA) / sum(ABBA + BABA), by = .(CHROM, START, END, ID1, ID2, ID3, REP, VARTYPE)]
dstat_info$all[, DSTAT := sum(ABBA - BABA) / sum(ABBA + BABA), by = .(ID1, ID2, ID3, REP, VARTYPE)]
```


Get missingness data

```{r getMissingness}
tmp <- list()
for (i in 1:5) {
    tmp[[i]] <- list()
    for (chunk in 1:16) {
        fn <- str_interp("results/missingness/merged.chunk${chunk}.mpileup.rep${i}.DP_${dstat_info$DP}.missingness.w_${dstat_info$wsize}.tsv")
        tmp[[i]][[chunk]] <- fread(fn, header = TRUE)
    }
    tmp[[i]] <- rbindlist(tmp[[i]])

    # Some windows span multiple chunks so sum those
    tmp[[i]][, CHROM := str_remove(str_remove(CHROM, "_RagTag$"), "^chrlg")]
    tmp[[i]][, ID := str_c(CHROM, "_", START)]
    tmp[[i]][, DUPE := duplicated(ID)]
    for (dup_x in tmp[[i]][, which(DUPE)]) {
        dup_id <- tmp[[i]][dup_x, ID]
        ori_i <- tmp[[i]][, which(ID == dup_id & !DUPE)[1]]
        for (cn in str_subset(names(tmp[[i]]), "^MISSING")) {
            x <- tmp[[i]][dup_x, get(cn)]
            tmp[[i]][ori_i, eval(cn) := get(cn) + x]
        }
    }
    tmp[[i]] <- tmp[[i]][!(DUPE)]
    tmp[[i]][, DUPE := NULL]
    tmp[[i]][, ID   := NULL]
    tmp[[i]][, REP  := i]
    tmp[[i]][, CHROM := factor(CHROM, levels = as.character(1:7))]
}

dstat_info[["miss"]] <- rbindlist(tmp)
rm(tmp)

# Get proportion for each window
j <- which(str_starts(names(dstat_info$miss), "MISSING"))
m <- as.matrix(dstat_info$miss[, ..j])
newvals <- rowSums(sweep(m, 2, 0:nrow(pop_d), `*`)) / (12 * rowSums(m))
dstat_info$miss[, MISSINGPROP := newvals]
```


## Plot

### Missingness across samples
```{r plotMissingDensity, fig.width = 5, fig.height = 4}
plt_d <- dstat_info$miss[, QuantileOverReplicates(MISSINGPROP, prefixes = "MISSINGPROP"), by = .(CHROM, START, END)]

p <- ggdensity(plt_d, x = "MISSINGPROP_med", fill = "grey",
               xlab = "Proportion missing GTs")
print(p)
```

### Summing across all windows
```{r plotDstatAll, fig.width = 5, fig.height = 4}
for (v in vartypes) {
    pts_d <- dstat_info$all[VARTYPE == v, QuantileOverReplicates(DSTAT, prefixes = "DSTAT"), by = .(ID1, ID2, ID3)]
    lines_d <- dstat_info$all[VARTYPE == v, .(MIN = min(DSTAT), MAX = max(DSTAT)), by = .(ID1, ID2)]

    use_ylim <- c(-1, 1) * lines_d[, max(abs(c(range(MAX), range(MIN))))]

    for (i in str_c("ID", 1:2)) {
        pts_d[, eval(i) := ShortenID(get(i))]
        lines_d[, eval(i) := ShortenID(get(i))]
    }

    pts_d[, REGION := "Allopatry\nwith cuspidata"]
    pts_d[str_starts(ID2, "D6"), REGION := "Sympatry\nwith cuspidata"]

    lines_d[, REGION := "Allopatry\nwith cuspidata"]
    lines_d[str_starts(ID2, "D6"), REGION := "Sympatry\nwith cuspidata"]

    dodge_w <- 0.8

    use_colors <- Load4SpPal("2drum")[c("drumBlue", "drumPink")]
    names(use_colors) <- str_c(c("Allopatry", "Sympatry"), "\nwith cuspidata")

    corner_text <- data.table(x = -Inf, y = Inf, label = v, ID1 = lines_d$ID1[1])

    p <- ggplot() +
        geom_linerange(aes(x = ID2, ymin = MIN, ymax = MAX, color = REGION), data = lines_d,
                       alpha = 0.4, linewidth = 10) +
        geom_linerange(aes(x = ID2, ymin = DSTAT_lwr, ymax = DSTAT_upr, color = REGION), data = pts_d,
                       position = position_dodge2(width = dodge_w)) +
        geom_point(aes(x = ID2, y = DSTAT_med, color = REGION), data = pts_d,
                   position = position_dodge2(width = dodge_w), shape = 1) +
        geom_hline(yintercept = 0, linetype = 2) +
        geom_text(aes(x = x, y = y, label = label), data = corner_text,
                  hjust = -0.05, vjust = 1.3, size = 7, color = "#444444", family = "Arial") +
        facet_wrap(vars(ID1), nrow = 1) +
        scale_color_manual(values = use_colors, name = "P. drummondii region") +
        ylim(use_ylim) +
        xlab("P. drummondii individual") +
        ylab("D-statistic") +
        theme(axis.text.x = element_text(angle = 30, hjust = 1, size = 8),
              strip.background = element_rect(fill = Load4SpPal()["roem"]))
    print(p)

    plist[[v]] <- p
    # ggsave(p, filename = "tmp.pdf", width = 5, height = 4)
}
```

### Correlation between Dstat and missingness

```{r corrMissDstat, fig.width = 8, fig.height = 8}
plt_d <- dstat_info$w[VARTYPE == "SNP", .(DSTAT = median(DSTAT)), by = .(ID1, ID2, CHROM, START)]
miss_d <- dstat_info$miss[, QuantileOverReplicates(MISSINGPROP, prefixes = "MISSINGPROP"), by = .(CHROM, START)]
plt_d <- merge(plt_d, miss_d, by = c("CHROM", "START"))
rm(miss_d)


p <- ggplot(plt_d, aes(x = MISSINGPROP_med, y = DSTAT)) + geom_bin2d() +
    scale_fill_viridis() +
    facet_grid(ID1 ~ ID2) +
    xlab("Proportion missing GTs") +
    ylab("Median D")

print(p)
# ggsave(p, file="tmp.pdf")
```

# Windowed D-statistic

Are there any shared windows across *all* quartets?

Plan, plot chr1 for a few quartets to show top-scoring D windows (250 kb)

(In the future, explore anti-correlated windows?)

We ignore windows with fewer than 1000 informative snps and with more than 50% missing genotypes


```{r dwindowIntersect}
min_inform_snps <- 1000
max_missingness <- 0.5

tmp <- dstat_info$miss[
    ,
    QuantileOverReplicates(MISSINGPROP, prefixes = "MISSINGPROP"),
    by = .(CHROM, START)
][
    MISSINGPROP_med < max_missingness
]

wnames <- tmp[, str_c(CHROM, "_", START)]

dat <- dstat_info$w[
    str_c(CHROM, "_", START) %in% wnames
][
    VARTYPE == "SNP"
][
    (BBAA + ABBA + BABA) > min_inform_snps,
    QuantileOverReplicates(DSTAT, BBAA, ABBA, BABA, prefixes = c("DSTAT", "BBAA", "ABBA", "BABA")),
    by = .(CHROM, START, END, ID1, ID2, ID3)
]


dat <- merge(dat, tmp, by = c("CHROM", "START"))
dat[, WINDOW := str_c(CHROM, "_", START)]

all_trios <- dat[, 1, by = .(ID1, ID2, ID3)]
all_trios[, V1 := NULL]
setorder(all_trios, ID1, ID2, ID3)
sd_cutoff <- 3

outlier_windows <- list()
for (sp in c("drum", "roem")) {
    outlier_windows[[sp]] <- list()
    for (i in seq_len(nrow(all_trios))) {
        ids <- all_trios[i, c(ID1, ID2, ID3)]
        if (sp == "drum") {
            outlier_windows[[sp]][[i]] <- dat[ID1 == ids[1]][ID2 == ids[2]][ID3 == ids[3]][as.vector(scale(DSTAT_med)) > sd_cutoff, WINDOW]
        } else {
            outlier_windows[[sp]][[i]] <- dat[ID1 == ids[1]][ID2 == ids[2]][ID3 == ids[3]][as.vector(scale(DSTAT_med)) < -sd_cutoff, WINDOW]
        }
    }
}
```

### Outlier windows shared across all comparisons
```{r outlierShared}
intersect_outliers <- list()
for (sp in names(outlier_windows)) {
    count_tab <- as.data.table(table(unlist(outlier_windows[[sp]])))
    setnames(count_tab, c("V1", "N"), c("WINDOW", "COUNT"))
    out_d <- dat[WINDOW %in% count_tab[["WINDOW"]], .(CHROM = CHROM[1],
                                                      START = START[1],
                                                      END = END[1],
                                                      MISSINGPROP = MISSINGPROP_med[1]), by = WINDOW]
    out_d <- merge(out_d, count_tab, by = "WINDOW")
    out_d[, WINDOW := NULL]
    setorderv(out_d, c("COUNT", "CHROM", "START"), order = c(-1, 1, 1))
    print(head(out_d))

    intersect_outliers[[sp]] <- out_d
    out_fn <- str_interp("results/outlier_windows.shared.${sp}.bed")
    fwrite(out_d, file = out_fn, col.names = FALSE, row.names = FALSE, quote = FALSE, sep = "\t")
}
```

### All pairwise comparisons

Compute all pairwise comparisons of distinct trios and the number of matching outlier windows

First let's get all the distinct pairs of trios that don't include any samples in common

```{r pwiseOutlierOverlap1}
all_pairs <- t(combn(seq_len(nrow(all_trios)), m = 2))
mask <- apply(all_pairs, 1, function(x) length(unique(unlist(all_trios[x]))) == 6)
uniq_pairs <- all_pairs[mask, ]
```

Next, compute number of overlaps in each pairwise comparison
```{r pwiseOutlierOverlap2}
overlap_d <- list()
for (sp in names(outlier_windows)) {
    overlap_d[[sp]] <- data.table(TRIO_A = uniq_pairs[, 1], TRIO_B = uniq_pairs[, 2], DIRECTION = sp)
    for (i in seq_len(nrow(uniq_pairs))) {
        trio_idx <- uniq_pairs[i, ]
        wA <- outlier_windows[[sp]][[trio_idx[1]]] # nolint: object_name_linter.
        wB <- outlier_windows[[sp]][[trio_idx[2]]] # nolint: object_name_linter.
        overlap_d[[sp]][i, N_OUTLIER_A := length(wA)]
        overlap_d[[sp]][i, N_OUTLIER_B := length(wB)]
        overlap_d[[sp]][i, N_OUTLIER_I := length(intersect(wA, wB))]
        overlap_d[[sp]][i, N_OUTLIER_U := length(union(wA, wB))]
    }
}
overlap_d <- rbindlist(overlap_d)
```

Plot overlap distribution

```{r pwiseOutlierOverlapPlot, fig.width = 5, fig.height = 4}
overlap_d[, FILL := str_c("cusp <=> ", DIRECTION)]
use_colors <- Load4SpPal()[c("drum", "roem")]
names(use_colors) <- str_c("cusp <=> ", names(use_colors))

p1 <- ggdensity(overlap_d, x = "N_OUTLIER_I", fill = "FILL") +
    scale_fill_manual(guide = "none", values = use_colors) +
    xlab("Number of overlapping outlier windows")

p2 <- ggboxplot(overlap_d, x = "FILL", y = "N_OUTLIER_I", fill = "FILL", orientation = "horizontal") +
    scale_fill_manual(name = "Introgression direction", values = use_colors) +
    theme(
        axis.title = element_blank(),
        axis.line = element_blank(),
        axis.text.x = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks = element_blank()
    )

p <- p2 / p1 + plot_layout(heights = c(1, 4))
print(p)
# ggsave(p, file = "tmp.pdf", width = 5, height = 4)
```

### Windows with Outliers (chr1)
```{r plotWindowsWithOutliers, fig.width = 12, fig.height = 3}
compare_trios <- c(19, 37)
tileheight <- 0.85

plt_d <- list()
for (i in compare_trios) {
    samps <- unlist(all_trios[i])
    lab <- str_c(ShortenID(samps), collapse = ",")
    plt_d[[lab]] <- dat[ID1 == samps[1] & ID2 == samps[2] & ID3 == samps[3]][CHROM == "1"]
    plt_d[[lab]][, LABEL := lab]
    plt_d[[lab]][, OUTLIER := WINDOW %in% c(outlier_windows$drum[[i]], outlier_windows$roem[[i]])]
}
plt_d <- rbindlist(plt_d)
plt_d[, COORD := START / 1e6]
plt_d[, LABEL := factor(LABEL)]

annot_d <- plt_d[(OUTLIER)]
annot_d[, SHAPE := as.integer(LABEL)]
int_w <- intersect(annot_d[SHAPE == 1, WINDOW], annot_d[SHAPE == 2, WINDOW])
annot_d[WINDOW %in% int_w, SHAPE := 3L]
annot_d[, SHAPE := as.character(SHAPE)]
annot_d[, COLOR := as.character(sign(DSTAT_med))]
annot_d[SHAPE == 1, Y := 1 - (tileheight / 2) * 1.1]
annot_d[SHAPE == 2, Y := 2 + (tileheight / 2) * 1.1]
annot_d[SHAPE == 3, Y := 1.5]

tilewidth <- 250e3 / 1e6

dmax <- plt_d[, max(abs(range(DSTAT_med)))]

use_gradient <- brewer.pal(11, "RdYlBu")
use_colors <- use_gradient[c(1, 11)]
names(use_colors) <- c("-1", "1")

use_shapes <- c(2, 6, 18)
names(use_shapes) <- as.character(1:3)

use_sizes <- c(1, 1, 2) * 1.4
names(use_sizes) <- names(use_shapes)

chr1_len <- 1129

limits <- list(x = c(-1, chr1_len + 1),
               y = range(annot_d$Y) + c(-0.1, 0.1))

box_d <- data.table(XMIN = rep(- tilewidth / 2, 2),
                    XMAX = rep(chr1_len + tilewidth / 2, 2),
                    YMIN = c(1, 2) - tileheight / 2,
                    YMAX = c(1, 2) + tileheight / 2)

p <- ggplot() +
    geom_rect(aes(xmin = XMIN, xmax = XMAX, ymin = YMIN, ymax = YMAX), data = box_d,
              color = "#777777", fill = NA, linewidth = 0.75) +
    geom_tile(aes(x = COORD, y = LABEL, fill = DSTAT_med), data = plt_d, width = tilewidth, height = tileheight) +
    scale_fill_gradientn(colors = use_gradient, limits = c(-dmax, dmax), name = "D-statistic") +
    geom_point(aes(x = COORD, y = Y, color = COLOR, shape = SHAPE, size = SHAPE), data = annot_d) +
    scale_color_manual(values = use_colors, guide = "none") +
    scale_shape_manual(values = use_shapes, guide = "none") +
    scale_size_manual(values = use_sizes, guide = "none") +
    xlab("Chr1 Position (Mb)") +
    coord_cartesian(xlim = limits$x, ylim = limits$y, expand = FALSE) +
    theme(axis.title.y = element_blank(),
          axis.line.y = element_blank())
print(p)

plist[["tile"]] <- p
```

### Compute likelihood of overlap w/ bootstrap
```{r bootstrapOverlap, fig.width = 4, fig.height = 3}
tmp_d <- list()
for (i in compare_trios) {
    nm <- as.character(i)
    samps <- unlist(all_trios[i])
    tmp_d[[nm]] <- dat[ID1 == samps[1] & ID2 == samps[2] & ID3 == samps[3]]

    tmp_d[[nm]][, INDEX := i]
    tmp_d[[nm]][, OUTLIER := "."]
    for (sp in c("drum", "roem")) {
        tmp_d[[nm]][WINDOW %in% outlier_windows[[sp]][[i]], OUTLIER := sp]
    }
}
tmp_d <- rbindlist(tmp_d)

tab <- tmp_d[, table(WINDOW)]
joint_w <- names(tab)[tab == 2]
tmp_d <- tmp_d[WINDOW %in% joint_w]
nwindows <- tmp_d[, .N / 2]
boot_params <- tmp_d[, as.list(table(OUTLIER)), by = INDEX][, .(drum, roem)]

noverlap <- tmp_d[OUTLIER != ".", sum(table(WINDOW) == 2), by = .(OUTLIER)][, sum(V1)]

BootstrapWindowOverlap <- function(nwindows, outlier1, outlier2, noverlap, nboot = 5000) {
    SampleFunc <- function() {
        samp1 <- sample.int(nwindows, size = sum(outlier1), replace = FALSE)
        samp2 <- sample.int(nwindows, size = sum(outlier2), replace = FALSE)

        noverlap <- sum(samp1[1:outlier1[1]] %in% samp2[1:outlier2[1]]) +
            sum(samp1[-(1:outlier1[1])] %in% samp2[-(1:outlier2[1])])

        noverlap
    }

    boot_replicates <- replicate(SampleFunc(), n = nboot)
    pval <- sum(boot_replicates > noverlap) / nboot
    return(list("replicates" = boot_replicates, "pval" = pval))
}

boot_res <- BootstrapWindowOverlap(nwindows, unlist(boot_params[1, ]), unlist(boot_params[2, ]), noverlap)

plt_d <- data.table(X = boot_res$replicates)

p <- gghistogram(plt_d, x = "X", fill = "#333333", color = NA, binwidth = 1) +
    geom_vline(xintercept = noverlap, color = "red") +
    xlab("Overlapping\noutlier windows")
print(p)

plist[["boot"]] <- p
```

### Joint plot (windows + bootstrap)

```{r plotJoint, fig.width = 14, fig.height = 3}
p <- plist$tile + plist$boot + plot_layout(widths = c(12, 2))
print(p)
ggsave(p, file = "tmp.pdf", width = 14, height = 3)
```

### SNP + SV window correlation

#### Calculate variant correlation
```{r calcVarCorrelation}
pts_d <- list()
for (v in vartypes) {
    pts_d[[v]] <- dstat_info$all[VARTYPE == v, QuantileOverReplicates(DSTAT, prefixes = v), by = .(ID1, ID2, ID3)]
}
pts_d <- merge(pts_d[[1]], pts_d[[2]], by = c("ID1", "ID2", "ID3"))

xvar <- "SNP_med"
yvar <- "SV_med"

ols_res <- RunOLS(pts_d, xvar = xvar, yvar = yvar)
print(summary(ols_res$model))
```

```{r plotVarCorrelation, fig.height = 4, fig.width = 4}

use_lab <- FormatR2PvalLab(unlist(ols_res[c("rsq", "pval")]))

p <- ggplot() +
    geom_point(aes(x = .data[[xvar]], y = .data[[yvar]]), data = pts_d, alpha = 0.8) +
    geom_ribbon(aes(x = .data[[xvar]], y = .data[[yvar]], ymin = LWR, ymax = UPR), data = ols_res$fitdat, fill = "#333333", alpha = 0.3) +
    geom_line(aes(x = .data[[xvar]], y = .data[[yvar]]), data = ols_res$fitdat, color = "blue") +
    geom_hline(yintercept = 0, linetype = 2) +
    geom_vline(xintercept = 0, linetype = 2) +
    annotate(geom = "text", x = -Inf, y = Inf, hjust = -0.1, vjust = 1.8, label = use_lab, parse = TRUE) +
    xlab("D-statistic (SNP)") +
    ylab("D-statistic (SV)")
print(p)
plist[["SV_SNP_corr"]] <- p
# ggsave(p, file = "tmp.pdf", width = 4, height = 4)
```

### Testing ...

```{r plotDstatTesting}
trios <- list()
cusp_ind <- pop_d[POP == "cusp", ID_LONG][3]
trios[[1]] <- c("proe_Hopkins_702-2-3-1", "pdrum_Hopkins_656a",      cusp_ind)
trios[[2]] <- c("proe_Hopkins_708_36_1",  "pdrum_Hopkins_706b-21-c", cusp_ind)

sub_d <- list()
for (i in 1:2) {
    sub_d[[i]] <- dstat_info$w[
        VARTYPE == "SNP"
    ][
        ((ID1 == trios[[i]][1]) & (ID2 == trios[[i]][2]) & (ID3 == trios[[i]][3]))
    ][
        (BBAA + ABBA + BABA) > 400,
        QuantileOverReplicates(DSTAT, prefixes = "DSTAT"),
        by = .(CHROM, START, END)
    ]
}

plt_d <- merge(sub_d[[1]], sub_d[[2]], by = c("CHROM", "START", "END"), suffixes = c("_X", "_Y"))

lm_model <- lm(DSTAT_med_Y ~ DSTAT_med_X, data = plt_d)

OrthoDist <- function(m, xvar, yvar) {
    line_params <- c(coef(m)[2], -1, coef(m)[1])
    x <- m$model[, xvar]
    y <- m$model[, yvar]

    sign(lm_model$residuals) * abs(line_params[1] * x + line_params[2] * y + line_params[3]) / sqrt(sum(line_params[1:2]^2))
}

plt_d[, ORTHODIST := OrthoDist(lm_model, "DSTAT_med_X", "DSTAT_med_Y")]

plt_d[, EXTREME := as.vector(abs(scale(ORTHODIST))) > 5]

p <- ggscatter(plt_d, x = "DSTAT_med_X", y = "DSTAT_med_Y", color = "EXTREME", alpha = 0.3) +
    geom_vline(xintercept = 0, linetype = 2) +
    geom_hline(yintercept = 0, linetype = 2) +
    geom_abline(intercept = 0, slope = 1, color = "gray", linetype = 2)
print(p)
dev.off()
```

# Figure 4 full plot

```{r fullPlot, fig.width = 12, fig.height = 11.93651}
dadi_dim <- c(557, 365)

scatter_w <- 199
w <- dadi_dim[1] + scatter_w
tile_w <- w * 11 / 14

# row_h <- cumsum(c(w / 3, dadi_dim[2], w / 14 * 2.5))
row_h <- c(w / 3, dadi_dim[2], w / 14 * 2.5)

print(row_h[length(row_h)] / w * 12)

r1_shorten <- 20 # Shorten row 1 width by this amount

# top left bottom right
layout <- c(
    # DefineArea(0, 0,
    #            w = r1_shorten, h = row_h[1]),
    DefineArea(r1_shorten, 0,
               w = (w - r1_shorten) / 2, h = row_h[1]),
    DefineArea(r1_shorten + (w - r1_shorten) / 2, 0,
               w = (w - r1_shorten) / 2, h = row_h[1]),

    DefineArea(0, row_h[1],
               w = scatter_w, h = row_h[2]),

    DefineArea(0, row_h[1] + row_h[2],
               w = tile_w, h = row_h[3]),
    DefineArea(tile_w, row_h[1] + row_h[2],
               w = w - tile_w, h = row_h[3])
)

pSNP <- plist$SNP # nolint: object_name_linter.
pSV <- plist$SV # nolint: object_name_linter.
pCorr <- plist$SV_SNP_corr # nolint: object_name_linter.
pTile <- plist$tile # nolint: object_name_linter.
pBoot <- plist$boot # nolint: object_name_linter.

p <- ((pSNP + pSV + plot_layout(guides = "collect")) + pCorr +
    (free(pTile + plot_layout(guides = "keep"), side = "l")) + free(pBoot, side = "l")) +
    plot_layout(design = layout)

ggsave(p, file = "figure4.pdf", width = 12, height = 12 / max(layout$r) * max(layout$b))
print(p)
```

# Random non-outlier windows

To provide a fair comparison against the outlier windows use randomly sampled non-outlier windows.
We apply a 25% quantile cutoff on the median BBAA count and also require that the window pass this threshold
in all 48 trios.

```{r sampleNonOutlier}
set.seed(3027)

bbaa_thresh <- quantile(dat$BBAA_med, 0.25)
print(bbaa_thresh)

nsamp <- 100

ignore_w <- union(unlist(outlier_windows$drum), unlist(outlier_windows$roem)) # Ignore outlier windows
w_pool <- dat[!(WINDOW %in% ignore_w)][BBAA_med > bbaa_thresh, .N == nrow(all_trios), by = WINDOW][(V1), WINDOW]
samp_windows <- sample(w_pool, size = nsamp, replace = FALSE)

out_d <- dat[WINDOW %in% samp_windows, .(CHROM[1], START[1], END[1], MISSINGPROP_med[1]), by = WINDOW]
out_d[, WINDOW := factor(WINDOW, levels = samp_windows)] # Order by sample
setorder(out_d, WINDOW)
out_d[, WINDOW := NULL]

out_fn <- str_interp("results/samp_windows.bed")
fwrite(out_d, file = out_fn, col.names = FALSE, row.names = FALSE, quote = FALSE, sep = "\t")
```

# Save for debugging
```{r saveData}
save.image("results/figure4.RData")
# load("results/figure4.dstat.RData")
```
